---
title: "03_FooDB_ManualMatch100"
author: "Stephanie Wilson"
date: "January 2023"
output: html_document
---

# Database Matching: FooDB to ASA24 Ingredient Descriptions
## Step 3: Manually Matching 100 random ASA food descriptions to FooDB

__Required Input Files__
  - *FooDB_V2_descripcleaned.csv* - Output from 02_FooDB_TextProcessing
  - *remap_descripcleaned.csv* - Output from 02_FooDB_TextProcessing, a smaller dataframe than the full asa descriptions
  
__Information__
This script manually matches 100 food descriptions from ASA dietary recall data to FooDB.

  1) Generates a simple random sample of 100 food descriptions of remapped asa descriptions.
  2) Manual comparisons aided by ingredient subcode matching and pattern searching.
  3) Create a function that will match ASA24 with manually curated FooDB codes
  4) Export data frame with matched info
     - Output: Manual_Match100.csv
  
__Output__
  - Manual_Match100.csv

```{r Load packages, include= FALSE, warning=FALSE, message=FALSE}
library(dplyr) #Version 1.0.9
library(readxl)
```


```{r Load Data, include=FALSE}
Food_V2 = read.csv("FooDB/Food_V2_descripcleaned.csv", header=TRUE)
asa = read.csv('data/remap_descripcleaned.csv', header=TRUE)
```

### 1) Generates a simple random sample of 100 food descriptions from dietary recalls.

```{r Set Seed}
#initial_seed = as.integer(Sys.time()) #Output: 1667858872
#substring(initial_seed, first = nchar(initial_seed)-4, last = nchar(initial_seed)) #pulls out last five numbers
set.seed(58872)
```

```{r Pull out 100 distinct ingredient codes}
random100 = asa %>% 
  distinct(Ingredient_description_y, .keep_all = TRUE) %>% 
  sample_n(size = 100)

#filter in columns for matching
random100 = random100[,(4:6)]

#Create new column row for entry number
random100$entry = rownames(random100)
```

### 2) Manual comparisons aided by ingredient subcode matching and pattern searching

```{r Filter in specific columns from Food_V2}
#'foodb_id', 'orig_food_id', 'orig_food_common_name', 'food_V2_ID', 'Ingredient_subcode_drop'
Food_V2 = Food_V2 %>% 
  select(c('food_id', 'orig_food_id', 'orig_food_common_name', 'food_V2_ID'))
```


```{r Check potential hits with orig_food_id}
#Convert orig_food_id to numeric. 
#Note: note all orig_food_id are ingredient subcodes, this was done for comparison purposes.
Food_V2$orig_food_id = as.numeric(Food_V2$orig_food_id)

#Merge files
hits = left_join(random100, Food_V2, by=c('Ingredient_subcode' ='orig_food_id'), keep=TRUE)

#Optional File for Comparisions, Write to csv
# write.csv(hits, 'hits.csv', row.names = FALSE)
```

```{r Checks}
#how many rows don't have code matches?
table(is.na(hits$orig_food_id)) #14 don't have matches and will need pattern searching

# What are duplicated entries? Will need to consolidate
duplicated =  hits %>% 
  group_by(Ingredient_subcode) %>% 
  filter(n()>1)  #Any subcode with more than 1 entry is filtered IN
```


```{r Examine NAs in hits file and pattern search}
Chicken = Food_V2 %>% filter(grepl('chicken', orig_food_common_name, ignore.case = TRUE))
Chicken

Turkey = Food_V2 %>% filter(grepl('turkey.*raw', orig_food_common_name, ignore.case = TRUE))
Turkey

Biscuit = Food_V2 %>% filter(grepl('biscuit', orig_food_common_name, ignore.case = TRUE))
Biscuit

Food_V2 %>% filter(grepl('celery.*raw|raw.*celery', orig_food_common_name, ignore.case = TRUE))
```


```{r}
#Compare and enter FooDB matches manually by unique id, 
#matches goes in order from 1 to 100
FooDB_ID_match = c(701, 8293, 7115, 9405, 1283, # 1-5
                   1454, 4500, 1636, 1444, 3778, 
                   2615, 283, 6741, 4765, 6868, #11-15
                   4948, 7863, 9014, 7785, 5521, 
                   1379, 2606, 8283, 7977, 697, #21-25
                   5222, 1091, 2252, 1101, 1764, 
                   958, 253, 4946, 5463, 904, #31-35
                   3068, 3518, 2503, 938, 5727, 
                   1423, 1684, 4825, 1148, 4298, #41-45
                   4851, 527, 1500, 3749, 3734, 
                   5043, 9013, 7629, 5660, 545, #51-55
                   5719, 6688, 4941, 7834, 2331, 
                   9016, 7396, 5638, 4670, 5741, #61-65
                   1165, 3744, 355, 602, 247, 
                   4770, 859, 665, 1400, 9435, #71 -75
                   4212, 4947, 8301, 1076, 5503, 
                   448, 9016, 851, 264, 9014, #81 - 85
                   8329, 5044, 941, 9014, 4759, 
                   2200, 7580, 5489, 8020, 780, #91-95
                   8012, 3747, 7116, 4793, 499)
```

### 3) Function that will match ASA24 with _manually_ curated FooDB codes
```{r Function to pair match data}
matched <- function(x, y, data_x, data_y){
  df = data.frame(matrix(ncol=8, nrow=1)) #create empty df
  colnames(df) = c('Ingredient_subcode', 'Ingredient_description', 
                            'Ingredient_weight_g', 'entry', 'foodb_id', 'orig_food_id', 
                            'orig_food_common_name', 'food_V2_ID')
  Uniquefood = x
  FooDB_ID_match = y
  random100 = data_x
  Food_V2 = data_y
  df[x, 1:8] = cbind(random100[x,], Food_V2[y,]) %>%
      data.frame() 
  return(df)
}
```

### 4) Export dataframe with matched info
```{r Create dataframe with 100 random ASA descriptions and their manual FooDB match}
Manual_Match100 = matched(x = row.names(random100), y = FooDB_ID_match, data_x = random100, data_y = Food_V2)

#Drop a repeated column and reorder the unique FooDB ID
Manual_Match100 = Manual_Match100 %>% 
  relocate(food_V2_ID, .after = entry)

# Export file
write.csv(Manual_Match100, 'data/Manual_Match100.csv', row.names = FALSE)
```






