---
title: "E1B_Polyphenol_Estimation_PhenolID_Milk"
author: "Stephanie Wilson"
date: "August 2023"
output:
  pdf_document: default
  html_document: default
---

# Polyphenol Estimation
## Step 1B: Polyphenols in Milk in FooDB

__Required Input Files__  
  - *milk.sdf* - SDF from Naveja email communications
  - *FooDB_phenols.csv* - Phenols pulled out of Compounds.csv

__Information__  
In August 2023, it was discovered there were discrepanccies with compounds in milk originating from a [2019 paper included in FooDB](https://doi.org/10.1021/acs.jafc.9b00204). Milk compounds and their concentrations were scraped off of Foodb.ca with Dr. Jose Naveja helping us re-identify polyphenolic compounds in milk. Identification followed methods outlined in his 2018 paper in FL100. 

Specifically, this script does the following:

  1) Pull out ID's from provided Naveja SDF (file of identified phenolic compounds in FooDB).
  2) Match phenol IDs from SDF to FoodB's Compound List for additional compound information
  3) Revise Units
  
__Outputs__  
  - *milk_scrape_polyphenols.csv* - milk polyphenols content data
  - *milk_scrape_phenols.csv* - milk phenols content data
  
```{r Load Packages, message=FALSE}
library(cinf) ; library(tidyverse)
```


```{r Load Data}
# Load SDF dataset provided by Dr. Naveja
SDF = read_sdf('FooDB/milk.sdf')

# Load FooDB data, accessed September 2022
# Use Content.csv to assess all possible matches
polyphenols = read.csv('FooDB/FooDB_phenols.csv')

# All compounds from milk scraped off of FooDB
# File that was sent to Jose 
scrape = read.csv('FooDB/milk_compound_data.csv')
```


### 1) Pull out information from provided Naveja SDF (file of identified phenolic compounds in FooDB).

Pull out compound identifying information and polyphenol status from props
```{r}
# Template for map usage _map(list, function, 'item name or position number')_
ID = SDF %>% 
  map_chr(., pluck, 'props', 'compound_public_id') %>%
  tibble() %>%
  rename('compound_public_id' = 1)

# Jose added an additional column to let us know if it was polyphenolic or phenolic
polyphenol_status = SDF %>% 
  map_chr(., pluck, 'props', 'Polyphenol') %>%
  tibble() %>%
  rename('polyphenol_status' = 1) %>%
  mutate(polyphenol_status = ifelse(polyphenol_status == '1', 'polyphenol', 'phenol'))

ID_merge = cbind(ID, polyphenol_status)
```


### 2) Match phenol IDs from SDF to FoodB's Compound List for additional compound information
```{r Pull out selected phenolic IDs from FooDB Compounds}
milk_phenols = right_join(scrape, ID_merge, by ='compound_public_id') %>%
  mutate(compound_name = ifelse(compound_public_id=='FDB023335',
                                "3,3',4'5-Tetrahydroxystilbene", compound_name))

#How many compounds are unquantified?
table(is.na(milk_phenols$Average))[2]/nrow(milk_phenols)*100

# How many are polyphenols
table(milk_phenols$polyphenol_status=='polyphenol')
```


### 3) Revise Units

Given:  
Concentration (g/L) = (Concentration in micromoles/L) × (Molar mass in g/mol)

Example:  
Concentration in micromoles/L (uM): 500 uM
  - Molar mass in g/mol: 650.7900 g/mol
  - Concentration (g/L) = (500 × 10^-6 mol/L) × 650.7900 g/mol
  - Concentration (g/L) = 0.0005 mol/L × 650.7900 g/mol
  - *Concentration (g/L) = 0.325395 g/L*
  
Given: 
1 g/L = 10 mg/100g

Example: 
  - Concentration (mg/100g) = 0.325395 g/L × 10 = 3.25395 mg/100g

```{r}
molar = milk_phenols %>%
  filter(units =='uM') %>%
  select(-units) %>%
  mutate(low_end = ((low_end * 10^-6) * mono_mass) * 10,
         high_end = ((high_end * 10^-6) * mono_mass) * 10,
         Average = ((Average * 10^-6) * mono_mass) * 10,
        units = 'mg/100 g')
```


```{r}
# What's missing concentration data?
unquantified = milk_phenols %>%
  filter(is.na(Average))

# Create a new version where everything's mg/100g
mg100g = milk_phenols %>%
  filter(units =='mg/100 g') %>%
  
  #Add the updated molar data
  full_join(molar) %>%
  
  # Add the missing values in too
  full_join(unquantified)
```

Export csv
```{r}
phenol_list = mg100g %>% filter(polyphenol_status == 'phenol')
polyphenol_list = mg100g %>% filter(polyphenol_status == 'polyphenol')

write.csv(phenol_list, 'FooDB/milk_scrape_phenols.csv', row.names = FALSE)
write.csv(polyphenol_list, 'FooDB/milk_scrape_polyphenols.csv', row.names = FALSE)
```

