---
title: "Q5_Polyphenol_Estimation_ASAMerge"
author: "Stephanie Wilson"
date: "September 2023"
output: html_document
---

#  Polyphenol Estimation
## Step 05: Merger with ASA

__Required Input Files__.                          
  - *codematched_final.csv* - Output from 08_FooDB_FNDDS_FullMatch_Part5    
  - *ingredientized_asa_10-2022.csv* - Output from 04_ingredientize_merge.rmd script, 
      - 'Ingredientized' Quality Controlled ASA24 Recall Data from FL100 Study.  
  - *FooDB_phenol_content.csv.bz2* - Output from E4_Polyphenol_Estimation_Units
  - *Food_updated.csv* - 01_FooDB_FooDBCleaning.ipynb
  - *FooDB_PE_RF.csv* - Output from E03_Polyphenol_Estimation_RF.Rmd
  - *FooDB_phenol_content_foodsums.csv* - Output from E4_Polyphenol_Estimation_Content

__Information__. 
This script merges ingredients within recall data to their ingredient equivalent in FooDB. The resulting output is subsequently merged with the FooDB polyphenol content file. 
    1) Merger of dietary recall data with FooDB matches and then polyphenol content     
      - This step also includes the creation of a new column, **pp_consumed**, which is the amount of polyphenols consumed in milligrams, which incorporates all possible retention factor data. 

__Output__. 
  - *FooDB/FooDB_phenol_content_ASAmerged.csv.bz2*

```{r message=FALSE}
library(tidyverse)
```


```{r Load Data}
#FooDB polyphenol quantities
FooDB_mg_100g = read.csv('FooDB/FooDB_phenol_content.csv.bz2') %>%
  distinct(food_id, compound_id, .keep_all = TRUE) %>%
  select(-c(food_name, orig_food_common_name)) %>%
  relocate(orig_content_avg, orig_unit_updated, .before = citation)

#Ingredientized ASA codes matched to FooDB
matched = read.csv('data/codematched_final.csv') %>%
  relocate('food_V2_ID', 'orig_food_common_name', 
                  .before = Ingredient_code) %>%
  select(-Ingredient_description)

#Ingredientized ASA
ingredients = read.csv('Ingredientize/data/ingredientized_asa_10-2022.csv') %>%
  select(-c(WWEIA.Category.description, UserID, EatWith)) %>%
  select(-(KCAL:D_CHEESE)) # remove ASA nutrient information

#Load updated FooDB Food ID's
food_sums = read.csv('FooDB/FooDB_phenol_content_foodsums.csv') 
```


### 1) Merger of dietary recall data with FooDB matches 
Merge is based on food_id. 

Link matched ASA/FooDB matched descriptions. Result is ingredientized ASA matched to FooDB.
```{r}
ASA = left_join(ingredients, matched, by = 'Ingredient_code') %>%
  relocate(Ingredient_code, Ingredient_description, food_id, food_V2_ID, 
           orig_food_common_name, mean_score, .after = Food_Description) %>%
  relocate(RecallNo, .after = UserName) %>%
  select(-c(food_V2_ID, orig_food_id, fndds_description, 
                   IntakeStartDateTime, IntakeEndDateTime, RecallAttempt))

message('Number of unique foods in ASA: ',
  length(unique(ASA$food_id)))
```

Merge FooDB-matched ASA Ingredient Codes with FooDB Polyphenol Content File.
  - Link between FooDB Polyphenol Content and ASA code-matched data is *food_id*.
```{r}
ASA_FooDB = ASA %>%
  dplyr::left_join(FooDB_mg_100g, by = 'food_id')
```
Check how many retention factors carried over.
```{r}
hasRF = ASA_FooDB %>% 
  dplyr::filter(!is.na(ASA_FooDB$aggregate_RF))

# How many compounds are quantified that have RF/#How many compounds are quantified
nrow(ASA_FooDB %>% 
       filter(!is.na(orig_content_avg_RFadj)) %>% 
                filter(!is.na(aggregate_RF)))/
nrow(ASA_FooDB %>% filter(!is.na(orig_content_avg_RFadj))) 
```
Summary
```{r}
message('Number of unique foods in ASA-FooDB (with RF): ',
  length(unique(hasRF$food_id)),
  
  '\nNumber of unique compounds in ASA-FooDB (with RF): ',
  length(unique(hasRF$compound_id)),
  
  '\nNumber of unique foods in ASA- FooDB (content is not zero, with RF): ',
  nrow(hasRF %>% dplyr::filter(orig_content_avg_RFadj > 0) %>% 
         dplyr::select(food_id) %>% distinct()),
  
  '\nNumber of unique compounds in ASA- FooDB (content is not zero, with RF): ',
  nrow(hasRF %>% dplyr::filter(orig_content_avg_RFadj > 0) %>%
    dplyr::select(compound_id) %>% distinct()))
```

### 2) Updating Content entries

Subset anything that wasn't originally quantified (Identified as 'Polyphenols Absent in FooDB' and  'Unquantified but Predicted or Expected' in Q2 Script). We will update these values and ensure we are also incorporating retention factors where applicable.
```{r}
#subset ids to be updated
id_updates = food_sums %>%
  dplyr::filter(!Status == 'Quantified') %>%
  dplyr::select(-c(Status, total_pp, Status_adj, food_public_id, food_name,
                   total_pp_adj))

content_updates = ASA_FooDB %>%
  dplyr::filter(food_id %in% id_updates$food_id) %>%
  dplyr::left_join(id_updates, by = 'food_id') %>%
  dplyr::select(-c(orig_content_avg_RFadj)) %>% #drop old values
  dplyr::rename('orig_content_avg_RFadj' = 'total_pp_RFadj') #replace with new values
```


```{r}
#subset status adjustment column to merge
status = food_sums %>%
  dplyr::select(c(food_id, food_name, Status_adj))

# create the new file with the adjusted contents
# Add status adjustment tag to the updated content file so we can trace what we've updated
ASA_FooDB_adj = ASA_FooDB %>%
  dplyr::filter(!food_id %in% id_updates$food_id) %>% #remove updated entries from original file
  dplyr::full_join(content_updates) %>% #Add updated entries back in
  dplyr::left_join(status, by = 'food_id') %>%
  dplyr::relocate(food_name, .after = food_id) %>%
  dplyr::select(-c(orig_content_avg)) #These don't have RF and new zeroing updates.
```


### 3) Adjusting Beverages

'Nectar' Beverages
Adjust for percent juice. Only a handful of users.
```{r}
#Nectar Food Codes
nectar_ids = c('64203020', '64204010')

nectar = ASA_FooDB_adj %>%
  dplyr::filter(FoodCode %in% nectar_ids) %>%
  dplyr::mutate(pp_consumed = ifelse(FoodCode == 64203020, (orig_content_avg_RFadj*0.01) *
           (Ingredient_consumed_g*(0.15)), (orig_content_avg_RFadj*0.01) *
           (Ingredient_consumed_g*(0.20))))
```


FooDB values for tea and coffee derived from Dr. Duke's Ethnobotanical database are provided as dry weight and require adjustment for brewing. The following ratios were used for adjustment:  
  - *Tea*: 1g dry weight per 50 mL water 
    -  [(Kowalsa et al 2021)](https://doi.org/10.3390/molecules26164773)
  - *Coffee*: 10 g dry weight per 180 mL water
    -  [SCA Golden Cup Standard, American](https://sca.coffee/research/protocols-best-practices)

Tea Adjustment    
```{r}
#FoodB food_ids for tea products
tea_food_ids = c(1021, 957, 38, 940, 849, 939, 517, 748)

tea = ASA_FooDB_adj %>%
  dplyr::filter(food_id %in% tea_food_ids) %>%
  dplyr::mutate(pp_consumed = ifelse(citation == 'DUKE', 
                                     (orig_content_avg_RFadj*0.01) * (Ingredient_consumed_g * (1/50)), # Adjustment
                                     (orig_content_avg_RFadj*0.01) * Ingredient_consumed_g))
```

Coffee Adjustment
```{r}
#FoodB food_ids for coffee products
coffee_food_ids = c(891, 60, 58, 59)

coffee = ASA_FooDB_adj %>%
  dplyr::filter(food_id %in% coffee_food_ids) %>%
  dplyr::mutate(pp_consumed = ifelse(citation == 'DUKE', 
                                     (orig_content_avg_RFadj*0.01) * (Ingredient_consumed_g * (10/180)), # Adjustment
                                     (orig_content_avg_RFadj*0.01) * Ingredient_consumed_g))
```

Replace coffee and tea values. Add pp_consumed column by multiplying amount of the polyphenol content (mg/100g multiply by 0.01 to get mg/g) by ingredient consumed (grams) to get the polyphenol amount consumed (mg).
```{r message = FALSE}
ASA_FooDB_updated = ASA_FooDB_adj %>%
  filter(!food_id %in% tea_food_ids) %>% #take out old values
  filter(!food_id %in% coffee_food_ids) %>% #take out old values
  filter(!FoodCode %in% nectar_ids) %>% #take out old values
  mutate(pp_consumed = (orig_content_avg_RFadj*0.01) * 
                  Ingredient_consumed_g) %>%
  full_join(tea) %>% #add new values
  full_join(coffee) %>% #add new values %>%
  full_join(nectar) %>%
  select(-Modified) %>%
  relocate(pp_consumed, orig_content_avg_RFadj, .before = orig_unit)
```

```{r EXPORT}
write.csv(ASA_FooDB_updated, 'FooDB/FooDB_phenol_content_ASAmerged.csv.bz2', 
          row.names = FALSE)
```

Since we created orig_content_avg, The file needs additional cleaning to remove the additional polyphenol entries 

