---
title: "E1C - Yogurt Corrections"
author: "Stephanie Wilson"
date: "September 2023"
output: html_document
---

__Overview__
After the appearance of yogurt in the top ten food contributors to total polyphenol intake (Before 9/15/23), the yogurt entry needs inspected. This scripts assessed what goes into yogurt and identified that soy yogurt entries were simultaneously included with food_id 634, yogurt and food_id 792, soy yogurt causing yogurt (634) to be inflated. 

__Purpose__
Incorrect yogurt entries are identified for removal and exported as a csv. 

__Input__
  - *Content.csv.* - FooDB data (version 1) downloaded Sept 2022 from FooDB.ca/downloads
  - *Food_updated.csv* - 01_FooDB_FooDBCleaning.ipynb
  
__Output__
  - *yogurt_soy_correction.csv* 

Load Data
```{r}
Content = read.csv("FoodB/Content.csv.bz2", header=TRUE) %>% 
  filter(!orig_food_common_name == 'Lipid from Arabidopsis (PathBank)') %>%
  filter(!food_id == 16420)

Food = read.csv('FooDB/Food_updated.csv') %>%
  select(c(id, name, public_id)) %>%
  dplyr::rename('food_id' = 'id',
         'food_name' = 'name',
         'food_public_id' = 'public_id')

Compound = read.csv('FooDB/Compound.csv.bz2') %>%
  select(c(id, public_id, name))
```

### 1) Generate unique lists of original food entries underlying each FooDB food_id
```{r}
distinct = Content %>%
  filter(source_type == 'Compound') %>%
  select(c(id, food_id, orig_food_common_name)) %>%
  # Add Food Names
  left_join(Food, by = 'food_id') %>%
  distinct(food_id, food_name, orig_food_common_name, .keep_all = TRUE) %>%
  arrange(food_id) %>%

  # Which original foods occur across multiple foods? Highlighted by a count >2
  mutate(count = table(orig_food_common_name)[orig_food_common_name])
```

### 2) Examine Yogurt, Isolate Errors
```{r}
yogurt = distinct %>%
  filter(food_id %in% c(634, 792)) %>%
  
  # Isolate Duplicates
  filter(count ==2) %>%
  
  # Remove Fructose as unrelate
  filter(!orig_food_common_name == 'Fructose')

yogurt
```

Where is the content data? Are they duplicated across 634 and 792?
```{r}
yogurt_compounds = Content %>%
  filter(source_type == 'Compound') %>%
  filter(food_id %in% c(634, 792)) %>%
  filter(orig_food_common_name %in% yogurt$orig_food_common_name) %>%
  #Add Compound Names
  left_join(Compound, by = c('source_id' = 'id')) %>%
  relocate(public_id, name, .after = source_id) %>% 
  left_join(Food) %>%
  select(id, source_id, public_id, name, source_type, food_id, orig_food_common_name,
         orig_content, orig_unit, citation) %>%
  arrange(orig_food_common_name, source_id)

yogurt_compounds
```

It does indeed look like content values across food_ids are the same, so we are okay to remove soy entries from 634. We will need the orig_food_common_name in order to remove this from the main content file

Export the file
```{r}
# Isolate the entries that need removed
yogurt_corrections = yogurt_compounds %>% 
  filter(!food_id == 792)

write.csv(yogurt_corrections, 'FooDB/yogurt_soy_correction.csv', row.names = FALSE)
```

