---
title: "E2_Polyphenol_Estimation_Content"
author: "Stephanie Wilson"
date: "`r Sys.Date()`"
output: html_document
---


# Polyphenol Estimation
## Step 2: Preparation of Phenol Content File

__Required Input Files__  
  - *FooDB_phenols.csv* - output from E1_Polyphenol_Estimation_FooDB_PhenolID.Rmd
  - *Food_V2_descripcleaned.csv* - output from 02_FooDB_TextProcessing.ipynb
  - *Content.csv.* - FooDB data (version 1) downloaded Sept 2022 from FooDB.ca/downloads
  - *Content_updated.csv* - output from 01_FooDB_FooDBCleaning.ipynb
  - *Food_updated.csv* - 01_FooDB_FooDBCleaning.ipynb
  - *FooDB/milk_scrape_polyphenols.csv* - Milk polyphenols from E1B_Polyphenol_Estimation_FooDB_PhenolID_Milk.Rmd
  - *yogurt_soy_correction.csv* - output form E1C_Yogurt_Correction.Rmd

__Information__  
This script prepares the Content file with cleaned text descriptions and merges this information with our phenol list. The output is ready to be matched with dietary recall data.

  1) Replace Content_updated food descriptions with cleaned text descriptions
  2) Add endogenous compounds back into Content_updated
  3) Consolidate Changes from Internal Milk Misclassifications
  4) Merge phenol compound data with Content_updated with cleaned food descriptions
     - Important Note: Content file needs to be filtered for source_type=='Compounds'
  5) Unit Check    

__Outputs__  
  - *FooDB_phenol_premerge.csv.bz2* - Pre-averaged content values for retention factor matching
  - *FooDB_polyphenols_V2.csv* - List of Polyphenols within FooDB.
  
```{r Load packages, message=FALSE}
library(tidyverse)
```


Load in Correction Files
```{r}
# Contains orig_food_common_name to take out from food_id 634
yogurt = read.csv('FooDB/yogurt_soy_correction.csv')

# Milk specific polyphenols
Milk = read.csv('FooDB/milk_scrape_polyphenols.csv') %>%
  rename('compound_public_id' = 1,
         'compound_name' = 2,
         'orig_min' = 'low_end',
         'orig_max' = 'high_end',
         'orig_content' = 'Average',
         'orig_unit' = 'units') %>%
  select(c(compound_public_id, compound_name, orig_min, orig_max, orig_content, 
           orig_unit, citation)) %>%
  mutate(source_type = 'Compounds', 
         food_id = 632,
         food_name = 'Milk (Cow)',
         food_public_id = 'FOOD00618')
```


```{r Load data}
# Load FooDB phenols
FooDB_phenols_raw = read.csv('FooDB/FooDB_phenols.csv', header=TRUE) %>%
  rename('smiles' = 'cas_number',
         'inchi' = 'moldb_inchikey',
         'mono_mass' = 'moldb_inchi',
         'inchikey' = 'moldb_smiles')

# Load cleaned text FooDB descriptions
Food_V2 = read.csv("FooDB/Food_V2_descripcleaned.csv", header=TRUE) %>%
  select(c(orig_food_common_name, food_V2_ID))

# Load FooDB data
# Original Content File needed for adding endogenous compounds back in
Content_updated = read.csv('FooDB/Content_updated.csv.bz2', header=TRUE)

Content = read.csv("FoodB/Content.csv.bz2", header=TRUE) %>% 
  filter(!orig_food_common_name == 'Lipid from Arabidopsis (PathBank)') %>%
  filter(!food_id == 16420)

# Load FooDB data, accessed September 2022
# Use Content.csv to assess all possible matches
Compound = read.csv('FooDB/Compound.csv.bz2') %>%
  select(c(id, public_id))

#Load updated FooDB Food ID's
Food = read.csv('FooDB/Food_updated.csv') %>%
  select(c(id, name, public_id)) %>%
  dplyr::rename('food_id' = 'id',
         'food_name' = 'name',
         'food_public_id' = 'public_id')
```

### 1) Replace food descriptions with cleaned text descriptions
```{r}
#match food_V2_ids, then replace old desription with new
Content_updated[["new_descrip"]] = Food_V2[match(Content_updated[['food_V2_ID']], Food_V2[['food_V2_ID']]), 'orig_food_common_name']

#delete old food descriptions, 
#rename new_id to orig_food_common_name
Content_updated = Content_updated %>% 
  select(-c('orig_food_common_name')) %>%
  dplyr::rename(orig_food_common_name = new_descrip) %>%
  relocate(c('orig_food_common_name', 'preparation_type', 
             'food_V2_ID'), .after = orig_food_id)
```


### 2) Add endogenous compounds back into Contents_updated to recreate Content file
We had previously removed Pathbank and endogenous compound (HMDB) entries from Content.csv in in 01_FooDB_FooDBCleaning.ipynb. Based on our exploration of phenol matches in E1_Polyphenol_Estimation_FooDB_PhenolID.Rmd, we know that none of the 3063 polyphenolic compounds have entries related to Pathbank. However, several polyphenolic compounds had an endogenous description, so we need to add endogenous compounds back in. We will also give them a food id, 9914.

Content first needs to be filtered for source_type=='Compounds' which is where you'll find polyphenols. 
  - source_type == 'Nutrients' retrieves 'classical nutrients' in dietary reporting
  
```{r message=FALSE}
#Identify the Endogenous Compounds
Endogenous = anti_join(Content, Content_updated, by = 'id') %>% 
  mutate(food_V2_ID = 9914)

# Add endogenous back into Content_updated
# We will keep out items from HMDB
Content_filtered  = full_join(Content_updated, Endogenous) %>% 
  filter(source_type == 'Compound') %>%
  left_join(Compound, by = c('source_id' = 'id')) %>%
  relocate(public_id, .after = id) %>%
  
  # Remove soy yogurt entries
  filter(!id %in% yogurt$id)
```


### 3) Consolidate Changes from Internal Milk Misclassifications

Create modification lists
```{r}
#What compounds did we not previously detect before? We need to add this to our polyphenol list
add = anti_join(Milk, FooDB_phenols_raw, by = c('compound_public_id' ='public_id'))

# Three compounds in milk were found to not match 
remove = data.frame(public_id = c('FDB097404', 'FDB097405', 'FDB097408'))
```


```{r}
# Here are some additional compounds previously detected that did not match with their structure
checklist = data.frame(public_id = c(
    "FDB097215", "FDB097260", "FDB097304", "FDB097323", "FDB097324",
    "FDB097334", "FDB097343", "FDB097345", "FDB097346", "FDB097349",
    "FDB097353", "FDB097354", "FDB097378", "FDB097381", "FDB097385"))

check = checklist %>% 
  left_join(FooDB_phenols_raw, by = 'public_id')  %>%
  left_join(Content_filtered, by = 'public_id') %>%
  
  #Let's look at the foods that are not milk
  filter(!food_id ==632)

# What are the unique food ids 
unique(check$public_id)
```

```{r}
# Are any of these missing quantification data? 
table(is.na(check$orig_content))

#Awesome. Let's go ahead and add these compounds to our removal list
remove_update = remove %>%
  full_join(checklist)

add_v2 = add %>%
  rename('name' = 'compound_name',
         'public_id' = 'compound_public_id') %>%
  select(name, public_id)

# Update the full polyphenol list
FooDB_phenols = FooDB_phenols_raw %>%
  full_join(add_v2) %>%
  filter(!public_id %in% remove_update$public_id)

write.csv(FooDB_phenols, 'FooDB/FooDB_polyphenols_V2.csv', row.names = FALSE)
```

### 4) Merge phenol compound data with Content Data with cleaned food descriptions

Let us merge and clean our file for export. 
```{r}
FooDB_phenol_content = Content_filtered %>%
  filter(public_id %in% FooDB_phenols$public_id) %>%
  left_join(FooDB_phenols[-1], by = 'public_id') %>%
  left_join(Food, by = 'food_id') %>%
  relocate('food_name', 'orig_food_part', 'food_public_id', .after = 'food_id') %>%
  #remove columns for brevity
  select(-c(moldb_iupac:subklass, id, orig_food_scientific_name,  
            orig_citation, creator_id, updater_id, created_at, 
            updated_at,  orig_unit_expression)) %>% 
  rename('compound_id' = 'source_id',
         'compound_name' = 'name',
         'compound_public_id' = 'public_id') %>%
  relocate(compound_id, compound_public_id, compound_name, 
           .before = orig_content) %>%
  arrange(food_id)
```


It appears a soy entry slipped through an earlier cleaning step. Quickly correct and replace in content file.
```{r}
soy = FooDB_phenol_content %>% 
  filter(food_id == 16420) %>% 
  mutate(food_id = 272,
         food_name = 'Other soy product',
         food_public_id = 'FOOD00271')

FooDB_phenol_content = FooDB_phenol_content %>%
  filter(!food_id == 16420) %>%
  add_row(soy)
```

Let us add in the corrected milk content adjustments.
```{r}
FooDB_content = FooDB_phenol_content %>%
  
  # Take out old entries
  filter(!food_id == 632) %>%
  
  #Add new entries
  full_join(Milk)
```


### 5) Unit Check
```{r}
table(FooDB_content$orig_unit)
```

```{r}
FooDB_content %>% filter(orig_unit =='mg/l')
```

SW Update:  
Manually cross-checked a random assortment of polyphenol entries with citation reference across several unit types. It appears 'orig_content' entries reflect mg/100 mg values and not units presented in orig_unit.
 - *FooDB Error*: where orig_unit == 'mg/kg', online FooDB.ca are in mg/g values but are reported as mg/100g.
  
Let us write our current version out so we can add in retention factors in another script.
```{r}
write.csv(FooDB_content, 'FooDB/FooDB_phenol_premerge.csv.bz2', row.names = FALSE)
```
