---
title: "Duke_Filtering"
author: "Stephanie Wilson"
date: "`r Sys.Date()`"
output: html_document
---

#  Filter Dr. Duke's Ethnobotanical Records

__Information__. 
It was noted that in Dr. Duke's polyphenol content entries that there are food parts that are measured but not commonly consumed. This script goes through Dr. Duke compounds (that have a food match with our dietary data) and filters in the food parts that are commonly consumed. We will update FooDB_phenol_premerge.csv.bz2 with these filtered entries downstream. 

__Required Input Files__.                          
  - *codematched_final.csv* - Output from 08_FooDB_FNDDS_FullMatch_Part5    
  - *ingredientized_asa_10-2022.csv* - Output from 04_ingredientize_merge.rmd script, 
      - 'Ingredientized' Quality Controlled ASA24 Recall Data from FL100 Study.  
  - *FooDB_phenol_content.csv.bz2* - Output from Q4_Polyphenol_Estimation_Units
 
__Output__. 
  - *Duke_Filtered_Content.csv* - Dr. Duke entries that have been cleaned for commonly consumed food parts.
  
```{r Load Data}
#FooDB polyphenol quantities
FooDB_phenol_content = read.csv('FooDB/FooDB_phenol_premerge.csv.bz2') %>%
  filter(!compound_public_id == 'FDB001111')

#Dietary Data
ingredients = read.csv('Ingredientize/data/ingredientized_asa_10-2022.csv') %>%
  select(-c(WWEIA.Category.description, UserID, EatWith)) %>%
  select(-(KCAL:D_CHEESE)) # remove ASA nutrient information


matched = read.csv('data/codematched_final.csv') %>%
  relocate('food_V2_ID', 'orig_food_common_name', 
                  .before = Ingredient_code) %>%
  select(-Ingredient_description)

# Join our dietary data with their corresponding FooDB food code
asa = left_join(ingredients, matched, by = 'Ingredient_code')
```

### Examine and Select Food parts 

Dr. Duke's Ethnobotanical Database contains multiple food parts, including those that are not commonly consumed. For each food, the most common food parts are filtered in.

```{r}
duke = FooDB_phenol_content %>%
  filter(citation == 'DUKE')%>%
  
  # Filter in food entries that have more than one food part so we can analyze further
  group_by(food_id) %>%
  filter(n_distinct(orig_food_part) >= 1) %>%
  ungroup()
```

Look at the specific combos
```{r}
part_combos = duke %>% 
  filter(food_id %in% asa$food_id) %>%
  select(food_id, food_name, orig_food_part) %>%
  distinct(food_id, orig_food_part, .keep_all = TRUE)

part_combos
```

Create a list of FooDB ids and for each id, the parts that we want.
```{r}
desired_orig_food_parts = data.frame(
  food_id = c(6, 9, 11, 12, 13, 13,
              14, 15, 15, 16, 17, 
              18, 18, 19, 19, 21,
              22, 26, 32, 37, 38, 38, 38, 
              39, 39, 40, 43, 43,
              47, 47, 49, 49, 49,
              54, 56, 56, 57, 57,
              59, 61, 61, 61, 61,
              63, 65, 67, 67, 68, 
              79, 81, 85, 85, 85,
              86, 88, 88, 92, 94,
              98, 101, 105, 105, 105, 105, 
              106, 115, 118, 119, 119, 
              121, 121, 122, 122, 124, 124,
              125, 129, 130, 131, 131, 131,
              133, 134, 135, 141, 141, 
              144, 147, 149, 150, 150,
              151, 151, 152, 153, 174,
              175, 175, 178, 178, 183, 183,
              200, 204, 204, 204, 205,
              206, 245, 444, 46, 50, 
              77, 95, 139, 158, 192), 
  desired_part = c("Bulb", "Leaf", "Seed", "Fruit", "Plant", "Leaf",
                   "Plant", "Plant", "Seed", "Seed", "Root",
                   "Root", "Plant", "Plant", "Leaf", "Shoot", 
                   "Seed", "Plant", "Sprout Seedling", "Seed", "Leaf", "Plant", "Stem", 
                   "Flower", "Plant", "Fruit", "Seed", "Leaf",
                   "Seed", "Plant", "Plant", "Leaf", "Root",
                   "Fruit", "Fruit", "Fruit Juice", "Fruit", "Fruit Juice",
                   "Seed", "Leaves", "Plant", "Fruit", "Shoot", 
                   "Silk Stigma Style", "Fruit", "Seed", "Leaf", "Rhizome",
                   "Seed", "Fruit", "Seed", "Seed Coat", "Plant",
                   "Seed", "Seed", "Plant", "Root", "Seed", 
                   'Seed', 'Seed', "Fruit", "Fruit Juice", "Fruit Epidermis", 'Pericarp',
                   'Fruit', "Leaf", "Seed", "Plant", "Leaf",
                   'Fruit', 'Pericarp', "Plant", "Leaf","Plant", "Leaf",
                   'Seed', "Plant", "Fruit", "Plant", "Leaf", "Stem",
                   "Seed", "Seed", "Fruit", "Seed", "Fruit",
                   "Fruit", "Fruit", "Fruit", "Fruit", "Pericarp",
                   "Fruit", "Fruit Juice", "Fruit Juice", "Root", "Fruit",
                   'Tuber', 'Tuber Epidermis', 'Plant', 'Leaf', "Plant", "Leaf",
                    "Seed", 'Exocarp', 'Fruit Juice', 'Fruit', 'Seed',
                   'Rhizome', "Root", "Fruit", "Plant", 'Bark',
                   'Leaf', 'Plant', 'Fruit', 'Shoot', 'Fruit'))
```


### Filter Duke Data to include desired food parts.
```{r}
result = duke %>%
  filter(food_id %in% desired_orig_food_parts$food_id & 
           orig_food_part %in% desired_orig_food_parts$desired_part)

write.csv(result, 'FooDB/Duke_Filtered_Content.csv', row.names = FALSE)
```


