---
title: "Q4_Polyphenol_Estimation_Content"
author: "Stephanie Wilson"
date: "September 2023"
output: html_document
---
 
# Polyphenol Estimation
## Step 4: Aggregation of Polyphenol Content

__Required Input Files__  
  - *FooDB_phenol_premerge.csv.bz2* - Output from E2_Polyphenol_Estimation_Content.Rmd
  - *Food_updated.csv* - 01_FooDB_FooDBCleaning.ipynb
  - *FooDB_PE_RF.csv* - contains FooDB/Phenol Explorer matches with retention factor data. Also includes those foods that did not match. Output from E3_Polyphenol_Estimation.
  - *Duke_Filtered_Content.csv* - Output from E3A_Polyphenol_Estimation_Duke_Filtering.Rmd

__Information__  
This script prepares the Content file with cleaned text descriptions and merges this information with our phenol list. Polyphenols are aggregated at the food level (not orig_food_common_name) to match what is shown on FooDB. The output is ready to be matched with dietary recall data.

  1) Replace Dr. Duke's Entries
  2) Data Preparation
  3) Review
  4) Polyphenol Totals and Updating the Content File

__Outputs__  
  - *FooDB_phenols_Content.csv.bz2* - Phenols pulled out of Compounds.csv and matched to FooDB's Compounds file with cleaned text descriptions.
  - *FooDB_phenol_content_foodsums.csv* - Summed polyphenol intake per unique food id in FooDB. Specific foods not present in FooDB or present but not quantified have had their concentrations adjusted. 
  
```{r Load packages, message=FALSE}
library(tidyverse)
```


```{r Load data}
# Remove Chinese Tannin as an outlier compound
FooDB_phenol_content_raw = read.csv('FooDB/FooDB_phenol_premerge.csv.bz2') %>%
  filter(!compound_public_id == 'FDB001111')

#Retention Factor Data
RF = read.csv('FooDB/FooDB_PE_RF.csv') %>%
  mutate(orig_food_id = as.character(orig_food_id),
         orig_source_id = as.character(orig_source_id))

Food = read.csv('FooDB/Food_updated.csv') %>%
  select(c(id, public_id, name)) %>%
  rename('food_id' = 'id',
         'food_public_id' = 'public_id',
         'food_name' = 'name')

# Filtered Dr. Duke's Content data
duke = read.csv('FooDB/Duke_Filtered_Content.csv') %>%
  mutate(orig_food_id = as.character(orig_food_id))
```


```{r}
message('Number of unique foods in FooDB content (unaggregated): ',
  length(unique(FooDB_phenol_content_raw$food_id)),
  
  '\nNumber of unique compounds in FooDB content (unaggregated): ',
  length(unique(FooDB_phenol_content_raw$compound_id)),
  
  '\nNumber of unique foods in FooDB content (quantified, unaggregated): ',
  nrow(FooDB_phenol_content_raw %>% dplyr::filter(!is.na(orig_content)) %>% select(food_id) %>% distinct()),
  
  '\nNumber of unique compounds in FooDB content (quantified, unaggregated): ',
  nrow(FooDB_phenol_content_raw %>% dplyr::filter(!is.na(orig_content)) %>% select(compound_id) %>% distinct()))
```

### 1) Replace Dr. Duke's Entries

```{r}
FooDB_phenol_content = FooDB_phenol_content_raw %>%
  
  #Take out the Duke citations
  filter(!citation == 'DUKE') %>%
  
  #Add the filtered Duke citations
  full_join(duke)
```



### 2) Data Preparation

**Aggregation for the FooDB polyphenol content**.   
  1) We will update orig_unit to reflect the unit presented in orig_content.
  2) Create average original content column to replicate values observed on [FooDB.ca](www.foodb.ca). 
  - *Note* - This action also moves up from the ingredient level to a broader food classification.
  - *Note* - This action removes individual entries that went into the average.

```{r}
#Merge with unit file
FooDB_phenol_content_updated = FooDB_phenol_content %>%
  #remove Phenol Explorer Entries
  dplyr::filter(!citation == 'PHENOL EXPLORER') %>% 
  
  #Add in our Phenol Explorer Entries that have been matched
  dplyr::full_join(RF) %>% 
  dplyr::mutate(orig_unit_updated = 'mg/100g') %>% 
  
  #Average
  dplyr::group_by(compound_public_id, food_id) %>%
  dplyr::mutate(orig_content_avg = mean(orig_content, na.rm = TRUE)) %>%
  dplyr::ungroup() %>% #NEW
  dplyr::relocate(orig_content_avg, .before = orig_unit) %>% 
  dplyr::relocate(food_public_id, .before = food_name) %>%
  dplyr::relocate(food_V2_ID, .before = orig_food_common_name) %>%
  dplyr::arrange(food_id, compound_id)%>%
  dplyr::distinct(food_id, compound_id, .keep_all = TRUE) %>% 
  
  #Removes ind. entries
  dplyr::select(-c(orig_food_id, orig_food_part, orig_content, orig_min,
                   orig_max, standard_content, export)) %>%
  dplyr::mutate(orig_content_avg_RFadj = ifelse(is.na(aggregate_RF),
                                                orig_content_avg,
                                       orig_content_avg*aggregate_RF)) %>%
  dplyr::relocate(aggregate_RF, orig_content_avg_RFadj, .after =  orig_content_avg)
```


```{r}
#Export Compound & Content Information
write.csv(FooDB_phenol_content_updated, 'FooDB/FooDB_phenol_content.csv.bz2', 
          row.names = FALSE)
```

### 3) Review
How many have quantified content values?
```{r}
message('Number of unique foods in FooDB content (aggregated): ',
  length(unique(FooDB_phenol_content_updated$food_id)),
  
  '\nNumber of unique compounds in FooDB content (aggregated): ',
  length(unique(FooDB_phenol_content_updated$compound_id)),
  
  '\nNumber of unique foods in FooDB content (quantified, aggregated): ',
  nrow(FooDB_phenol_content_updated %>% dplyr::filter(!is.na(orig_content_avg)) %>% select(food_id) %>% distinct()),
  
  '\nNumber of unique compounds in FooDB content (quantified, aggregated): ',
  nrow(FooDB_phenol_content_updated %>% dplyr::filter(!is.na(orig_content_avg)) %>% select(compound_id) %>% distinct()))
```

Great. This should match the numbers before aggregation as file was aggregated according to food_id and compound_id.

How many have quantified retention factors?
```{r}
message('Number of unique foods in FooDB content (aggregated): ',
  length(unique(FooDB_phenol_content_updated$food_id)),
  
  '\nNumber of unique compounds in FooDB content (aggregated): ',
  length(unique(FooDB_phenol_content_updated$compound_id)),
  
  '\nNumber of unique foods in FooDB content (with RF, aggregated): ',
  nrow(FooDB_phenol_content_updated %>% dplyr::filter(!is.na(aggregate_RF)) %>%
         select(food_id) %>% distinct()),
  
  '\nNumber of unique compounds in FooDB content (with RF, aggregated): ',
  nrow(FooDB_phenol_content_updated %>% dplyr::filter(!is.na(aggregate_RF)) %>%
         select(compound_id) %>% distinct()))
```

```{r}
#How many entries are quantified?
quantified = FooDB_phenol_content_updated %>%
  dplyr::filter(!is.nan(orig_content_avg))

# How many of the quantified entries have 
table(is.na(quantified$aggregate_RF))

564/nrow(quantified)*100
```

### 4) Polyphenol Totals and Updating the Content File

Create dataframe with polyphenol totals for each food_id within FooDB. *Total_pp* and *Total_pp_adj* units are mg/100g
```{r}
Food_Sums = FooDB_phenol_content_updated %>%
  drop_na(food_id) %>%
  dplyr::group_by(food_id) %>%
  dplyr::mutate(total_pp = sum(orig_content_avg, na.rm = TRUE),
         total_pp_RFadj = sum(orig_content_avg_RFadj, na.rm = TRUE)) %>%
  dplyr::distinct(food_id, .keep_all = TRUE) %>%
  dplyr::select(c(food_id, food_public_id, food_name, total_pp, total_pp_RFadj)) %>%
  dplyr::arrange(desc(total_pp))

#Stray Entry originating from Content file that needs updated
num899 = Food_Sums %>% filter(food_id == 899) %>%
  mutate(food_name = 'meats')
```


Quantified, Unquantified, Predicted Column Creation
```{r}
Food_Sums_Full_List = Food_Sums %>%
  dplyr::ungroup() %>%
  dplyr::filter(!food_id ==899) %>%
  dplyr::add_row(num899) 

Food_insert = Food %>%
  dplyr::anti_join(Food_Sums_Full_List, by = 'food_id')

Food_Sums_Full_List = Food_Sums_Full_List %>%
  dplyr::full_join(Food_insert) %>%
  dplyr::mutate(Status = case_when(
    total_pp > 0 ~ 'Quantified',
    total_pp == 0 ~ 'Unquantified but Predicted or Expected')) %>%
  dplyr::mutate(Status = coalesce(Status, 'Polyphenols Absent in FooDB')) #replace NAs
```


Identify meat and seafood foods from 'Unquantified but Predicted or Expected'. Zero them and reclassify them as quantified. 
```{r}
#Identify meat and seafoods 
meat_seafood = data.frame(food_id = c(334, 505, 549, 399, 293, 483, 541, 
                                       535, 590, 358, 373, 280, 303, 310,
                                       316, 319, 317, 354, 364, 365, 378,
                                       383, 430, 435, 438, 451, 491, 540, 
                                       542, 555, 554, 595, 611, 624, 625),
                           total_pp_adj = 0,
                           Status_adj = 'Quantified') %>%
  dplyr::left_join(Food_Sums_Full_List, by = 'food_id') %>%
  dplyr::relocate(total_pp_adj, Status_adj, .after = Status)
```


Reclassify 'Polyphenols Absent in FooDB' as zeros.  
Reclassify 'Unquantified but Predicted or Expected' as NAs, except for the meat and seafood foods which have already been re-quantified as zeros.  
  - per Danielle's recommendation, email 2/10/23
```{r}
#These foods have no linked polyphenols in FooDB
Absent_adjusted = Food_Sums_Full_List %>% 
  dplyr::filter(Status == 'Polyphenols Absent in FooDB') %>%
  dplyr::mutate(total_pp_adj = 0,
                total_pp_RFadj = 0,
         Status_adj = 'Quantified')

# These foods do have linked polyphenols in FooDB but there's no available concentration
Unquantified_Adjusted = Food_Sums_Full_List %>% 
  dplyr::filter(Status == 'Unquantified but Predicted or Expected') %>%
  dplyr::filter(!food_id %in% meat_seafood$food_id) %>%
  dplyr::mutate(Status_adj = 'Unquantified but Predicted or Expected',
                total_pp_adj = as.numeric('NA'),
                total_pp_RFadj = as.numeric('NA'))
```

Merge edits and Export.
```{r}
Food_Sums_Full_List_final = Food_Sums_Full_List %>%
  filter(Status == 'Quantified') %>%
  mutate(total_pp_adj = total_pp,
         Status_adj = 'Quantified') %>%
  full_join(meat_seafood) %>%
  full_join(Absent_adjusted) %>%
  full_join(Unquantified_Adjusted) 

write.csv(Food_Sums_Full_List_final, 'FooDB/FooDB_phenol_content_foodsums.csv',
          row.names = FALSE)
```


