---
title: "04_ingredientize_merge"
output: html_document
date: "2022-09-29"
---


# Ingredientize ASA24 food records
### Step 4: Creation of full ASA Ingredientization File

__Required Input Files__  
  - *asa_fndds_matched.csv* - From 02_ingredientize_mixedfoods output
  - *ingred_recode_remapped10202022.csv* - From 03_ingredientize_code_remap output
  
## Information  
This script does the following:

  1) Cleans the ASA/FNDDS + remap merge
  2) Add cleaned merged data to ASA data that did not need remapping
  3) Create Ingredient Consumed Total in ASA24 

__Output__    
  - *data/ingredientized_asa_10-2022.csv* - Ingredientized ASA data from FL100 Study
  

```{r Load Packages, message=FALSE}
library(tidyverse)
```

```{r LOAD DATA}
# Load combined ASA24 and FNDDS data
# Filter in specific columns to view data
asa = read.csv('data/asa_fndds_matched.csv') %>%
  rename('Ingredient_code' = 'Ingredient.code')


# Load remapped food_ingredient code data
code_remap = read.csv('data/ingred_recode_remapped10202022.csv') %>%
  select(-Ingredient_description_x)
```


### 1) Clean Recode File

```{r MERGE DATA FILES}
asa_recode = full_join(asa, code_remap, by = 'Ingredient_code')
```


```{r Clean Recode File}
# Create a list of hits from our remapped codes
matches = as.list(code_remap$Ingredient_code)

# Filter in Ingredient Codes that are in our code_remap file
# drop specific rows
# rename rows
asa_recode = asa_recode %>%
  filter(Ingredient_code %in% matches) %>%
  select(-c(Ingredient_code, Ingredient.description, Ingredient.weight..g., 
           Ingredient_weight_g_x)) %>%
  rename('Ingredient_code' = 'Ingredient_subcode', 
         'Ingredient_description' = 'Ingredient_description_y',
         'Ingredient_weight_g' = 'Ingredient_weight_g_y')
```


### 2) Filter and Merge in Ingredient Codes that did not need to be ingredientized
```{r FILTER}
asa =  asa %>%
  rename('Ingredient_description' = 'Ingredient.description',
         'Ingredient_weight_g' = 'Ingredient.weight..g.') %>%
  filter(Ingredient_code <10000000) 
```

```{r MERGE}
asa_all = full_join(asa, asa_recode)
```



### 3) Create Ingredient Consumed Total in ASA24

Generate amount consumed (g) for 'ingredientized' food codes. First, calculate 'Ingredient weight total' (in grams) to be used to generate proportion of Ingredient:
$Proportion\:of\:Ingredient = (Ingredient\:weight\:(g))/(Ingredient\:weight\:total)$. 

where Ingredient weight total is the sum of ingredient weights (in grams) for each remapped Ingredient code. This calculation is input into the following:      
$Ingredient\:consumed\:(g) = FoodAmt\: * Proportion \: of\:Ingredient$

```{r}
#Calculate Ingredient weight totals
totals = asa_all %>% 
  distinct(FoodCode, Ingredient_code, .keep_all = TRUE) %>%
  group_by(FoodCode) %>%
  summarise(Ingredient_weight_total_g = sum(Ingredient_weight_g))

#Merge the Ingredient Totals
# Create 'Proportion_ingredient' column
# Create 'Ingredient_consumed_g' column
asa_all = left_join(asa_all, totals) %>%
  mutate(Proportion_ingredient = Ingredient_weight_g/Ingredient_weight_total_g) %>%
  mutate(Ingredient_consumed_g = FoodAmt * Proportion_ingredient) %>%
  relocate(Ingredient_weight_total_g, Proportion_ingredient, 
           Ingredient_consumed_g, .after = Ingredient_weight_g)
```
Note: asa_all should join only by FoodCode. If it is joining by FoodCode and Ingredient_weight_g, then rerun section '2) Filter and Merge in Ingredient Codes that did not need to be ingredientized'.


```{r EXPORT}
write.csv(asa_all, 'data/ingredientized_asa_10-2022.csv', row.names = FALSE)
```

