---
title: "Prep File: Merge Study Variables"
author: "Stephanie Wilson"
date: "April 2023"
output: html_document
---

# Merge FL100 files

## Information
This script pools multiple FL100 files into one joint file. Files include data on demographics, SES, fecal markers, systemic markers, ASA24 HEI scores. Files also include summaries of total kcal

## Inputs. 
  - Binning Data
  - ASA24 Data
  - HEI
  - ASA24 Seasons
  - Stool 
  - SES
  - CRP 
  
## Outputs
- **FL100_merged_variables.csv**
- **FL100_merged_variables_FECAL.csv** - Individuals with stool samples over 24hr old and individuals without complete GI data have been removed
- **FL100_merged_variables_BLOOD.csv** - Individuals without complete blood parameters have been removed. 

```{r}
library(tidyverse); library(Hmisc) #for cut2
```

### Binning Data
```{r}
bindata = read.csv('data/FL100_Binning_Data.csv') %>%
  rename('UserName' = 'Participant.ID',
         'Bin' = 'Sampling.Bin.Number.',
         'BMI' = 'Final.BMI') %>%
  mutate(Sex = factor(Sex)) %>%
  select(c(UserName, Age, Sex, BMI, Age.Category, BMI.Category))
```

### Dietary Data

```{r}
#ASA data
asa = read.csv('data/fl100_recalls_qcd.csv')

# ASA Component and Total HEI Scores
hei = read.csv('data/CTSC24532USDAWHNRCNu-HEIscoreWithEthnicit_DATA_2023-05-31_1343.csv') %>%
  select(c(subject_id:hei_asa24_totalscore)) %>% 
  rename('UserName' = 'subject_id')
```


Additional calculation of total kcal and total fiber from ASA24 data.
```{r}
asa_averages = asa %>%
  group_by(UserName, RecallNo) %>%
  select(c(UserName, RecallNo, KCAL, FIBE, PROT)) %>%
  mutate(total_kcal = sum(KCAL, na.rm =TRUE),
         total_fiber = sum(FIBE, na.rm =TRUE),
         total_PRO = sum(PROT, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(UserName) %>%
  mutate(avg_total_kcal = mean(total_kcal, na.rm = TRUE),
         avg_total_fiber = mean(total_fiber, na.rm = TRUE),
         avg_total_PRO = mean(total_PRO, na.rm = TRUE)) %>%
  distinct(UserName, .keep_all = TRUE) %>%
  select(c(UserName, avg_total_kcal, avg_total_fiber, avg_total_PRO)) 
```


#### Stool
```{r}
stool = read.delim('data/FL100_stool_variables.txt', header = TRUE, sep = "\t", dec = ".") %>%
  rename('UserName' = 'subject_id')
```

#### CRP Data
```{r}
CRP = read.csv('data/CTSC24532USDAWHNRCNu-CRPData_DATA_LABELS_2023-05-30_1348.csv')%>%
  rename('UserName' = 1,
         'CRP_BD1' = 2,
         'CRP_BD3' = 3,
         'CRP_BD4' = 4)
```
        
### SES data
```{r}
ses = read.csv('data/FL100_SES.csv') %>%
  rename('UserName' = 'subject_id',
    'White' = 'race_ethnicity___1',
    'Hispanic_Latinx' = 'race_ethnicity___2',
    'Black' = 'race_ethnicity___3',
    'Asian' = 'race_ethnicity___4',
    'East_Asian' = 'race_ethnicity___5',
    'South_Asian' = 'race_ethnicity___6',
    'Southeast_Asian' = 'race_ethnicity___7',
    'Middle_Eastern' = 'race_ethnicity___8',
    'American_Indian_Alaska_Native' = 'race_ethnicity___9',
    'Native_Hawaiian_Pacific_Islander' = 'race_ethnicity___10',
    'Race_Declined' = 'race_ethnicity___11',
    'Race_Other' = 'race_ethnicity___12') %>%
  rowwise()%>%
  mutate(sum = sum(across(White:Native_Hawaiian_Pacific_Islander))) %>%
  select(c(UserName:other_ethnicity, edu_level, hhincome, sum)) 
```


```{r}
#Let us convert the binary data to more meaningful text descriptions
race_step1 = ses %>%
  mutate(
    White = ifelse(White == 1, 'White', ''),
    Hispanic_Latinx = ifelse(Hispanic_Latinx == 1, 'Hispanic or Latinx', ''),
    Black = ifelse(Black == 1, 'Black', ''),
    Asian = ifelse(Asian == 1, 'Asian', ''),
    East_Asian = ifelse(East_Asian == 1, 'East Asian', ''),
    South_Asian = ifelse(South_Asian == 1, 'South Asian', ''),
    Southeast_Asian  = ifelse(Southeast_Asian == 1, 'Southeast Asian', ''),
    Middle_Eastern = ifelse(Middle_Eastern == 1, 'Middle Eastern', ''),
    American_Indian_Alaska_Native = ifelse(
      American_Indian_Alaska_Native == 1, 'American Indian or Alaska Native', ''),
    Native_Hawaiian_Pacific_Islander = ifelse(
      Native_Hawaiian_Pacific_Islander == 1, 'Native Hawaiian or Pacific Islander', ''),
    Race_Declined = ifelse(Race_Declined == 1, 'Race Declined', '')) %>%
  select(c(UserName, White:other_ethnicity, sum))
```

Cleaning and Merging Racial Categories
```{r}
declined = race_step1 %>%
  filter(Race_Declined == 'Race Declined') %>%
  mutate(Race = Race_Declined) %>%
  select(c(UserName, Race))

multiracial = race_step1 %>%
  filter(sum > 1) %>%
  mutate(Race_description = paste(White, Hispanic_Latinx, Black, Asian, East_Asian,
                      South_Asian, Southeast_Asian, Middle_Eastern,
                      American_Indian_Alaska_Native,
                      Native_Hawaiian_Pacific_Islander, sep = ',  ')) %>%
  mutate(Race_description = gsub(x = Race_description, 
                                 replacement = ' ', pattern = ', {2,4}'),
         Race_description = gsub(x = Race_description, pattern = ", $", 
                                 replacement = ''), 
         Race_description = gsub(x = Race_description, pattern = "^, ", 
                                 replacement = ''),
         Race_description =gsub(x = Race_description, pattern = "\\s{1,6}",
                                replacement = ' ')) %>%
  mutate(Race = 'Two or More Races') %>%
  select(c(UserName, Race, Race_description))

other = race_step1 %>%
  filter(sum == 0) %>%
  filter(!UserName %in% declined$UserName) %>% #manual review, self-reported as multiracial
  mutate(Race = 'Two or More Races',
         Race_description = other_ethnicity) %>%
  select(c(UserName, Race, Race_description))

single = race_step1 %>%
  filter(!UserName %in% c(declined$UserName, multiracial$UserName, other$UserName)) %>%
  mutate(Race = paste(White, Hispanic_Latinx, Black, Asian, East_Asian, South_Asian,
                      Southeast_Asian, Middle_Eastern, American_Indian_Alaska_Native,
                      Native_Hawaiian_Pacific_Islander, sep = ', ')) %>%
  mutate(Race = gsub(x = Race, replacement = '', pattern = ', ')) %>%
  select(c(UserName, Race))

race = single %>%
  full_join(declined) %>%
  full_join(multiracial) %>%
  full_join(other) %>%
  arrange(UserName)
```

Update SES file
```{r}
ses_update = ses %>%
  select(-c(White:other_ethnicity, dov1, residence_city:birth_city, sum, age, sex, bmi)) %>%
  left_join(race, by = 'UserName')
```

### MERGE VARIABLES
```{r}
merged = bindata %>%
  left_join(ses_update, by = "UserName") %>%
  left_join(asa_averages, by = "UserName") %>%
  left_join(hei, by = 'UserName') %>%
  left_join(stool, by = "UserName") %>%
  left_join(CRP, by = 'UserName')

write.csv(merged, 'data/FL100_merged_variables.csv', row.names = FALSE)
```


Let us narrow this down further
```{r}
merged_succinct = merged %>%
  select(c(UserName,
           Age, Sex, BMI, #Main covariates + Sampling Scheme
           CRP_BD1, plasma_lbp_bd1, #Blood outcomes
           fecal_calprotectin, fecal_mpo, fecal_neopterin, #stool outcomes
           avg_total_fiber, hei_asa24_totalscore, #dietary predictors
           hei_asa24_totalveg:hei_asa24_addsug, #hei component scores
           bristol_num, fecal_ph, hhincome, edu_level, #additional predictors
           avg_total_kcal, After24h)) #for filtering steps, remove after filtering
```


```{r message = FALSE}
fecal = merged_succinct %>%
  filter(!is.na(avg_total_kcal)) %>% #removes 43 from incomplete ASA24
  filter(!After24h == 1) %>% #removes 30 individuals 
  filter(!is.na(fecal_neopterin)) %>% #removes 13
  select(-c(avg_total_kcal, After24h))

#Establish quartiles for each outcome
fecal$neo_quartile = as.numeric(cut2(fecal$fecal_neopterin, g = 4))
fecal$mpo_quartile = as.numeric(cut2(fecal$fecal_mpo, g = 4))
fecal$cal_quartile = as.numeric(cut2(fecal$fecal_calprotectin, g = 4))

write.csv(fecal, 'data/FL100_merged_variables_FECAL.csv', row.names = FALSE)
```


```{r}
blood = merged_succinct %>%
  filter(!is.na(avg_total_kcal)) %>% #removes 43 from incomplete ASA24
  filter(!is.na(CRP_BD1)) %>% # removes 2
  select(-c(avg_total_kcal, After24h))

#Establish quartiles
blood$CRP_quartile = as.numeric(cut2(blood$CRP_BD1, g = 4))
blood$LBP_quartile = as.numeric(cut2(blood$plasma_lbp_bd1, g = 4))

write.csv(blood, 'data/FL100_merged_variables_BLOOD.csv', row.names = FALSE)
```



