---
title: "Prep File: Compound Taxonomy Scraping"
author: "Stephanie Wilson"
date: "August 2023"
output: html_document
---

# Complete Taxonomy for Polyphenol Compounds

__Required Input Files__  
  - *FooDB_polyphenols_V2.csv* - output from Q2_Polyphenol_Quantification_Content.Rmd

__Information__  
This script scrapes the FooDB website to pull chemical compound classification data for polyphenols in the original data set that have missing taxonomy data. Specifically, this script:

  1) Extracts X-paths for Compound Classifications for each Uncategorized Polyphenol
  2) Run function on all uncategorized polyphenols and format output
  3) Update the original File

__Outputs__  
  - *FooDB_phenols_taxonomy.csv* - Our 3063 polyphenols with improved chemical compound classification data.

```{r LOAD PACKAGES, messages = FALSE}
library(dplyr); library(tidyr); library(purrr)
library(rvest); library(xml2); library(stringr)
source('Functions/get_taxonomy.R')
```

```{r LOAD DATA}
#Create urls
FooDB_phenols = read.csv('FooDB/FooDB_polyphenols_V2.csv') %>%
  mutate(compound_urls = paste0("https://foodb.ca/compounds/", public_id))

#Subset those that are unclassified
unclassified_compound_urls = FooDB_phenols %>%
  filter(kingdom == '') %>%
  select(public_id, compound_urls) 
```


### 1) Extract X-paths for Compound Classifications for each Uncategorized Polyphenol

Use an example compound to discover path. Message output is long.
```{r}
testurl = paste0("https://foodb.ca/compounds/FDB021899") #unclassified_compound_urls[150, 2]
doc = read_xml(testurl)
message(as.character(doc))
```

Test Xpaths
  - Return the character vector from an HTML element using XPath expression
```{r}
# Taxonomic Information
kingdom = xml_text(html_elements(doc, xpath = '//kingdom'))
superklass = xml_text(html_elements(doc, xpath = '//super_class'))
klass = xml_text(html_elements(doc, xpath = '//class'))
subklass = xml_text(html_elements(doc, xpath = '//sub_class'))
directparent =xml_text(html_elements(doc, xpath = '//direct_parent'))

# Create Dataframe
data = data.frame(kingdom, superklass, klass, subklass, directparent)
```

### 2) Run get_taxonomy to obtain classification on FooDB compounds

Test the function with a few compounds
```{r}
testurls = unclassified_compound_urls[1:2, 2]
output = sapply(testurls, get_taxonomy)
```


Run function on all uncategorized polyphenols
```{r}
output = sapply(unclassified_compound_urls$url_search, get_taxonomy)
```

Format Output
```{r}
#Transpose matrix and convert into dataframe
output.df = as.data.frame(t(output))

#Remove url base to leave just the compound id
rownames(output.df) = str_remove(rownames(output.df), "https://foodb.ca/compounds/") 

# Move row names to a column
output.df = tibble::rownames_to_column(output.df, 'compound_public_id')
```


Combine classification data
```{r}
#Simply and polish compound identifying information
names = FooDB_phenols %>%
  select(public_id, name) %>%
  rename('compound_public_id' = 'public_id',
         'compound_name' = 'name')

output.df.names = output.df %>%
  left_join(names, by = 'compound_public_id') %>%
  relocate(compound_name, .after = 'compound_public_id')
```


Adjust column structures for merger
```{r}
output.df.names = as.data.frame(apply(output.df.names, 2, as.character)) %>%
  mutate(kingdom = as.character(kingdom))
```


### 3) Update the original file

Remove rows that we searched for, and replace with updated data.
```{r}
FooDB_phenols_updated = FooDB_phenols %>%
  filter(kingdom == 'Organic compounds') %>%
  select(c('public_id', 'name', 'kingdom', 'superklass', 
           'klass', 'subklass')) %>%
  rename('compound_public_id' = 'public_id', 
         'compound_name' = 'name') %>%
  full_join(output.df.names) %>%
  mutate(kingdom = ifelse(kingdom == 'character(0)', 
                          'Unclassified', kingdom)) %>%
  mutate(superklass = ifelse(superklass == 'character(0)', 
                             'Unclassified', superklass)) %>%
  mutate(klass = ifelse(klass == 'character(0)', 
                        'Unclassified', klass)) %>%
  mutate(subklass = ifelse(subklass == 'character(0)', 
                           'Unclassified', subklass)) %>%
  mutate(directparent = ifelse(directparent == 'character(0)', 
                               'Unclassified', directparent))
```


```{r}
write.csv(FooDB_phenols_updated, 'FooDB/FooDB_phenols_taxonomy.csv', row.names = FALSE)
```
