---
title: "Prep File: Polyphenol Exploratory"
author: "Stephanie Wilson"
date: "March 2023"
output: html_document
---

#  Initial Exploration of polyphenol intakes.

__Required Input Files__.                          
  - *FooDB_phenol_content_ASAmerged.csv.bz2* - Output from Q5_Polyphenol_Estimation_ASAMerge. File still has compound level data.
  - *FooDB_phenols_taxonomy.csv* - Output from A0A_Scrape_Taxonomy.Rmd
  
__Information__. 
This script begins to explore polyphenol intakes in the FL100 study.

__Output__. 
  - *FL100_PP_Raw.csv* - Polyphenol content for each compound in foods listed in each recall test for each participant.
  - *FL100_PP_DailyAveraged.csv* - Polyphenol content for each compound in foods, averaged across each participant's recalls.
  - *FL100_Ingredient_Level_PP_Averages.csv* - Average polyphenol content for each ingredient.

```{r Load packages, message=FALSE}
library(tidyverse); library(hrbrthemes); library(ggrepel); library(wesanderson); library(readxl)
```

```{r Load Data}
ASA_FooDB = read.csv('FooDB/FooDB_phenol_content_ASAmerged.csv.bz2') %>%
  filter(!Ingredient_code == '999328') # Remove Vitamin D
PP_taxonomy = read.csv('FooDB/FooDB_phenols_taxonomy.csv') 
```


### 1) Exploratory DFs Creation

Create a simplified data frame showing the amount of each polyphenol consumed per food and individual. 
  - distinct() removes compound duplicates since we're using averages (what FooDB.ca uses)
  
```{r}
test = ASA_FooDB %>%
  filter(UserName == '8019') %>%
  distinct(RecallNo, food_id, compound_public_id, .keep_all = TRUE)

#total_daily_intake = total daily intake (mg) for each participant's recall
test_intake = test %>%
  group_by(RecallNo) %>%
  mutate(total_daily_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, total_daily_intake_mg)) %>%
  distinct(UserName, RecallNo, .keep_all = TRUE)

# Average for UserName 8019
mean(test_intake$total_daily_intake_mg) 
```
  
```{r}
exploratory = ASA_FooDB %>%
  distinct(UserName, RecallNo, food_id, compound_public_id, .keep_all = TRUE) %>%
  select(UserName, RecallNo, Food_Description, Ingredient_code, 
         Ingredient_description, food_id, food_name, compound_id, 
         compound_public_id, compound_name, Ingredient_consumed_g, 
         orig_content_avg_RFadj, pp_consumed, citation, Status_adj) %>%
  arrange(desc(pp_consumed))

write.csv(exploratory, 'data/FL100_PP_Raw.csv', row.names = FALSE)
```


But first, how many compounds are quantified and from how many foods? 
```{r}
quantified = exploratory %>% 
  filter(!orig_content_avg_RFadj == 'NA') 

# Summaries
message('Number of quantified distinct foods: ', 
        length(unique(quantified$food_id)))
message('Number of quantified distinct compounds: ', 
        length(unique(quantified$compound_id)))
message('Number of quantified & unquantified distinct foods: ', 
         length(unique(exploratory$food_id)))
message('Number of quantified & unquantified distinct compounds: ', 
         length(unique(exploratory$compound_id)))
```


Average across recalls for each compound in each reported food.
```{r}
exploratory_daily_average = exploratory %>%
  group_by(UserName, food_id, compound_id) %>%
  mutate(mean_daily_intake_mg = mean(pp_consumed, na.rm = TRUE)) %>%
  add_count(food_id, name = 'food_freq') %>%
  distinct(UserName, food_id, food_id, compound_id, .keep_all = TRUE)

write.csv(exploratory_daily_average, 'data/FL100_PP_DailyAveraged.csv',
          row.names = FALSE)
```

### 2) Exploratory Basics
```{r message = FALSE}
#Mean intake by polyphenol
mean_pp = exploratory %>%
  group_by(compound_id) %>%
  mutate(mean_intake_mg = mean(pp_consumed, na.rm = TRUE)) %>%
  select(c(compound_id, compound_name, mean_intake_mg)) %>%
  distinct(compound_id, compound_name, .keep_all = TRUE) 

#Number of unique food sources per polyphenol in FL100 cohort
unique_food_sources = exploratory %>%
  group_by(compound_name, compound_id) %>%
  distinct(food_name, .keep_all = TRUE) %>%
  count(sort = TRUE, name = 'num_uniq_foods') 

top_pp = left_join(mean_pp, unique_food_sources) %>%
  arrange(desc(mean_intake_mg), sort = TRUE)
```


```{r warning = FALSE}
uniquesourcesplot = ggplot(top_pp, aes(x = num_uniq_foods, y = mean_intake_mg,
                                    label = compound_name)) + 
  geom_point() +   
  theme_ipsum() + 
  labs(subtitle = 'Average Intake and Unique Food Sources for FoodB polyphenols',
       y='Mean Intake (mg)', 
       x='No. of Unique Food Sources') +
  geom_text_repel(data = subset(top_pp, mean_intake_mg > 350),
                  aes(label = compound_name), direction = 'x') +
  geom_text_repel(data = subset(top_pp, num_uniq_foods > 125),
                  aes(label = compound_name), direction = 'both', 
                  box.padding = unit(1.2, "lines"))
uniquesourcesplot
ggsave(plot = uniquesourcesplot, filename = 'images/unique_sources_compounds.png', width = 5, height = 4, units = "in", dpi = 300)
```


