---
title: "Analysis: Mappable Proportion"
author: "Stephanie Wilson"
date: "March 2023"
output: html_document
---


__Required Input Files__.                          
  - *FL100_PP_Raw.csv* - Polyphenol content for each compound in food, for each recall. Output from A0B_Exploration.Rmd.
  - *FL100_PP_DailyAveraged.csv* - Polyphenol content for each compound in food, averaged across recalls. Output from A0B_Exploration.Rmd.
  
__Information__. 
This script calculates and visualizes the mappable proportion of foods from ASA24 to FooDB.

__Output__. 
  - *FL100_Food_Polyphenol_Sum.csv* - Total polyphenol content for each food, for each recall. 
  - *Mappable_Percentage_Distribution.jpg*
  

```{r Load packages, message = FALSE}
library(tidyverse); library(Rfit); library(Hmisc); library(corrplot); library(ggpubr)
```


```{r Load Data}
PP_Raw = read.csv('data/FL100_PP_Raw.csv') 
PP_Daily = read.csv('data/FL100_PP_DailyAveraged.csv') 
```

### 1) Dataframe creation

Polyphenol Intake per food per recall
Polyphenol Intake per FooDB food *per recall* and the number of unique FooDB food IDs (not ASA Ingredient Codes) per each Recall for each participant. 
  - No compound clarity. Compounds have been summed.
  
```{r warning = FALSE}
pp_food_intake = PP_Raw %>%
  group_by(UserName, RecallNo, Ingredient_code) %>%
  mutate(food_pp_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>%
  distinct(Ingredient_code, .keep_all = TRUE) %>%
  ungroup() %>%
  group_by(UserName, RecallNo) %>%
  mutate(Num_Foods = n()) %>%
  select(-c(compound_id, compound_public_id, compound_name, 
            )) %>% #orig_content_avg, Ingredient_consumed_g, pp_consumed
  arrange(UserName, RecallNo)

#Adjust the food_pp_intake_mg entries that should be zero
corrections = pp_food_intake %>%
  filter(is.na(pp_consumed)) %>%
  mutate(food_pp_intake_mg = 'NA') %>%
  mutate(food_pp_intake_mg = as.numeric(food_pp_intake_mg))

#Fix the corrections in the main file
pp_food_intake = pp_food_intake %>%
  filter(!is.na(pp_consumed)) %>%
  full_join(corrections) 
```


```{r}
write.csv(pp_food_intake, 'data/FL100_Food_Polyphenol_Sum.csv', row.names = FALSE)
```

Ensure correct numbers with UserName 5001 as an example. 
```{r}
User5001 = PP_Raw %>%
  filter(UserName == '5001') %>%
  group_by(RecallNo, food_id) %>%
  mutate(food_pp_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>%
  distinct(food_id, .keep_all = TRUE) %>%
  ungroup() %>%
  group_by(RecallNo) %>%
  mutate(Num_Foods = n()) %>%
  select(-c(compound_id, compound_public_id, compound_name))
```


### 2) Assess percentage of foods mapped


#### 2a) Mappable Percentage Calculation
  - Num_Foods = number of foods in a recall
  - Num_Quantified_Foods = how many were quantified
  - percent_mapped = Num_Quantified_Foods/Num_Foods*100
  
```{r}
#count of Quantified Foods
counts = pp_food_intake %>%
  group_by(UserName, RecallNo) %>%
  count(is.na(food_pp_intake_mg), name = 'Num_Quantified_Foods')%>%
  select(-'is.na(food_pp_intake_mg)')

# add the counts to the main dataframe
pp_food_intake_counts = pp_food_intake %>%
  left_join(counts) %>%
  distinct(UserName, RecallNo, Ingredient_code, .keep_all = TRUE)

percentages = pp_food_intake_counts %>% 
  group_by(UserName, RecallNo) %>%
  mutate(percent_mapped = Num_Quantified_Foods/Num_Foods*100) %>%
  select(UserName, RecallNo, Num_Quantified_Foods, Num_Foods, percent_mapped) %>%
  distinct(UserName, RecallNo, .keep_all = TRUE) %>%
  mutate(Level = case_when(
    percent_mapped >= 90 ~ '>= 90%',
    percent_mapped >= 80 ~ '>= 80%',
    percent_mapped >= 70 ~ '>= 70%',
    percent_mapped >= 60 ~ '>= 60%',
    percent_mapped >= 50 ~ '>= 50%',
    percent_mapped >= 40 ~ '>= 40%',
    percent_mapped >= 30 ~ '>= 30%',
    percent_mapped >= 20 ~ '>= 20%',
    percent_mapped >= 10 ~ '>= 10%',
  ))
```

Determining distinct users
```{r}
test = percentages %>% 
  group_by(UserName) %>% 
  mutate(mean_percent_mapped = mean(percent_mapped)) %>%
  distinct(UserName, .keep_all = TRUE) %>%
  filter(mean_percent_mapped >= 80)

length(unique(test$UserName))

distinctusers = tibble(level = c('>= 90%', '>= 80%', '>= 70%', 
                                 '>= 60%', '>= 50%', '>= 40%'),
                           Num_Users = c(332, 350, 0, 0, 0, 0))
```

```{r}
mean(test$mean_percent_mapped)
sd(test$mean_percent_mapped)
```

```{r}
#isolate mean percent mapped
mean_percent_mapped = test %>% 
  select(UserName, mean_percent_mapped)

#merge the mean into the percentage df
percentages_updated = percentages %>%
  left_join(mean_percent_mapped) %>%
  select(-Level)

write.csv(percentages_updated, 'data/FL100_Mappable_Proportions.csv', 
          row.names = FALSE)
```


#### 2b) Plot percentages
```{r}
histogram = ggplot(percentages, aes(percent_mapped)) +
  geom_histogram() + 
  theme_bw() + 
  facet_grid(.~RecallNo) + 
  scale_x_continuous(limits = c(0, 110), breaks = seq(0, 100, 10)) + 
  labs(title = 'Percentage of Foods with Polyphenols Quantified to FooDB',
       subtitle = 'Distribution, Faceted by Recall Number', 
       x = 'Percent', y = 'Count')

barplot = ggplot(distinctusers, aes(x = level, y = Num_Users)) + 
  geom_bar(stat = 'identity', fill = 'darkgray') + 
  theme_bw() + 
  geom_text(aes(label = Num_Users), hjust = 1.3) + 
  labs(title = 'Number of Distinct Users above Mappable Percentage Thresholds',
       subtitle = 'Mappable Percentage, Average by User', 
       x = 'Mapped Threshold', y = 'User Count') + 
  coord_flip()

ggarrange(histogram, barplot, nrow =2, heights = c(1.2, 0.8))
```


Histogram of Average Mappable Proportion
```{r}
percentages_distinct = percentages %>%
  ungroup() %>%
  group_by(UserName) %>%
  mutate(Usermean = mean(percent_mapped)) %>%
  ungroup() %>%
  distinct(UserName, .keep_all = TRUE) %>%
  
  #Get Group mean
  mutate(mean = mean(Usermean),
         sd = sd(Usermean))

histogram_avg = ggplot(percentages_distinct, aes(percent_mapped)) +
  geom_histogram(color = "#000000", fill = "#4B0082") + 
  theme_bw() +
  geom_vline(aes(xintercept=mean, color="pink")) +
  geom_text(aes(x = mean + 3, y = 60, label = paste('Mean: ', signif(mean, digits = 3), '%'))) + 
  geom_text(aes(x = mean + 3, y = 55, label = paste('SD: ', signif(sd, digits = 3), '%'))) +
  labs(title = 'Distribution of Users',
       x = 'Mappable Foods (%)', 
       y = 'Count') +
  theme(legend.position = 'none')
histogram_avg

ggsave(plot = histogram_avg, filename = 'images/Mappable_Percentage_Distribution.jpg', height = 4, width = 6, unit = 'in')
```




