---
title: "Total Polyphenol Intake Analysis and Food Contributors"
author: "Stephanie Wilson"
date: "March 2023"
output: html_document
---

# Total Polyphenol Intake Analysis and Food Contributors

__Required Input Files__.                          
  - *FL100_PP_Raw.csv* - Polyphenol content for each compound in foods listed in each recall test. Output from A0B_Exploration.Rmd.
  - *FL100_PP_DailyAveraged.csv* - Polyphenol content for each compound in foods, averaged across each participant's recalls. Output from A0B_Exploration.Rmd.
  - *FL100_Food_Polyphenol_Sum.csv* - Total polyphenol content for each food, for each recall.
  - *FooDB_phenols_taxonomy.csv* - Taxonomic classifications of polyphenols. Output from A0A_Scrape_Taxonomy.Rmd.
  
__Information__. 
This script focuses on total polyphenol intake, specifically:
  1) Examines relationship between Energy-Adjusted Total Polyphenol Intake and Age, Sex, BMI, and Total HEI score
  2) Food Contributors of Total Polyphenol intake (Publication Table)
  3) Compound Class Contributors of Polyphenol Intake.

__Output__. 
 - Visual and Table: Top 10 Food Contributors of Total Polyphenol intake 

```{r Load packages, message = FALSE}
library(tidyverse); library(Rfit); library(Hmisc); library(corrplot); library(ggpubr); library(ggrain); library(viridis); library(ggridges); library(gt)
```


```{r Load Data}
#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list)

PP_Daily = read.csv('data/FL100_PP_DailyAveraged.csv') %>%
  filter(!compound_name %in% removal_list)

#multiple files including ASA averages for kcal and fiber
merged = read.csv('data/FL100_merged_variables.csv') %>%
  mutate(Sex = factor(Sex))

tax = read.csv('FooDB/FooDB_phenols_taxonomy.csv')
```


### 1) Do total daily intakes vary by Age, Sex, or BMI?

Sum polyphenol intake (mg) consumed for each recall, for each person, then averaged across recalls.
 - No compound or food clarity.
```{r}
#total_daily_intake = total daily intake (mg) for each participant's recall
total_daily_intake = PP_Raw %>%
  group_by(UserName, RecallNo) %>%
  mutate(total_daily_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, total_daily_intake_mg)) %>%
  ungroup() %>%
  group_by(UserName) %>%
  mutate(Avg_Total_Daily_PP_Intake_mg = mean(total_daily_intake_mg, na.rm = TRUE)) %>%
  distinct(UserName, .keep_all = TRUE) %>%
  select(-c(RecallNo, total_daily_intake_mg)) %>%
  left_join(merged, by = 'UserName') %>%
  arrange(UserName)  %>%
  ungroup()

totalkcal = total_daily_intake %>%
  select(c(UserName, avg_total_kcal))
```

Summary statistics of Covariates
```{r}
covariates = total_daily_intake %>%
  select(c(Age, Sex, BMI))

summary(covariates)
sd(covariates$Age)
sd(covariates$BMI)
```


### 1) Examine total polyphenol intake

```{r}
# Age, BMI, Sex only model
covariates.rfit = rfit(Avg_Total_Daily_PP_Intake_mg/(avg_total_kcal/1000) ~ Age + BMI + Sex, data = total_daily_intake)
summary(covariates.rfit)
```

```{r}
# With fiber and HEI total score
covariates.rfit.HEI = rfit(Avg_Total_Daily_PP_Intake_mg/(avg_total_kcal/1000) ~ Age + BMI + Sex + avg_total_fiber + hei_asa24_totalscore, data = total_daily_intake)
summary(covariates.rfit.HEI)
```


### 2) Food Contributors

```{r}
#Obtaining polyphenol intake for the whole cohort
overall_food_contributors = PP_Raw %>%
  left_join(totalkcal) %>%
  group_by(UserName, RecallNo, food_id) %>%
  mutate(food_ppintake = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, food_id, food_name, food_ppintake, avg_total_kcal)) %>%
  ungroup() %>%
  
  #Averages across recalls for each food
  group_by(UserName, food_id) %>% 
  mutate(Avg_food_ppintake = mean(food_ppintake/(avg_total_kcal/1000), na.rm = TRUE)) %>%
  distinct(UserName, food_id, .keep_all = TRUE) %>%
  ungroup() %>%
  
  #get overall food average
  group_by(food_id) %>% 
  add_count(food_id) %>% #get food frequency
  mutate(overall_avg_food_ppintake = mean(Avg_food_ppintake, na.rm = TRUE)) %>%
  select(c(food_id, food_name, overall_avg_food_ppintake, n)) %>%
  distinct(food_id, .keep_all = TRUE) %>%
  rename('food_freq' = 'n') %>%
  ungroup()

write.csv(overall_food_contributors, 'output/Ingredient_Intake_Frequency.csv', row.names = FALSE)
```


Obtaining the Average Ingredient Intake
```{r}
overall_ingredient_intake = PP_Raw %>%
  
  #Get all of the Ingredient sources (remove redundant information from polyphenol data)
  #Important to include Food_Description
  distinct(UserName, RecallNo, Food_Description, Ingredient_code, .keep_all = TRUE) %>%

  #Sum the amount of ingredient consumed per each particpant's recall
  # Here we pivot to the FooDB food id level instead of FNDDS codes
  group_by(UserName, RecallNo, food_id) %>%
  mutate(Ingredient_recalltotal = sum(Ingredient_consumed_g, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, food_id, food_name, Ingredient_consumed_g,
           Ingredient_recalltotal)) %>%
  ungroup() %>%
  
  #Average ingredient consumed across recalls for each participant
  group_by(UserName, food_id) %>% 
  mutate(Ingredient_UserAvgTotal = mean(Ingredient_recalltotal, na.rm = TRUE)) %>%
  distinct(UserName, food_id, .keep_all = TRUE) %>%
  ungroup() %>%
  
  #get overall food average ACROSS USERS  
  group_by(food_id) %>% 
  mutate(Ingredient_OverallAvgTotal = mean(Ingredient_UserAvgTotal, na.rm = TRUE)) %>%
  select(c(food_id, food_name, Ingredient_OverallAvgTotal)) %>%
  distinct(food_id, .keep_all = TRUE) %>%
  ungroup()
```


Begin filtering based off quartile designations to understand the top contributors.
```{r}
top_contributors = overall_food_contributors %>%
  left_join(overall_ingredient_intake) %>%
  mutate(freq_quartile = as.numeric(cut2(food_freq, g = 4)),
         intake_quartile = as.numeric(cut2(overall_avg_food_ppintake, g = 4))) %>%
  filter(freq_quartile == 4 & intake_quartile == 4) %>%
  arrange(desc(overall_avg_food_ppintake)) %>%
  slice_head(n = 10) %>%
  mutate(rank = ifelse(overall_avg_food_ppintake == max(overall_avg_food_ppintake), 1, rank(-overall_avg_food_ppintake)),
         grouping = ifelse(rank <=5, '1 - 5', '6 - 10'))
```


Table 1
```{r}
top_contributors_table_format = top_contributors %>%
  select(c(rank, food_name, food_freq, overall_avg_food_ppintake,
           Ingredient_OverallAvgTotal)) %>%
  mutate(overall_avg_food_ppintake = signif(overall_avg_food_ppintake, digits = 3),
         Ingredient_OverallAvgTotal = signif(Ingredient_OverallAvgTotal, digits = 3)) %>%
  rename('Rank' = 1,
         'Food' = 2,
         'User Frequency' = 3,
         'Polyphenol Intake (mg)' = 4,
         'Food Amount Consumed (g)' = 5)

write.csv(top_contributors_table_format, 'output/top_foodcontributors_table.csv', row.names = FALSE)

gt_tbl = gt(top_contributors_table_format)
gt_tbl
```


### 3) CLASS SUMMARY
```{r message = FALSE}
class_analysis = PP_Raw %>%
  left_join(tax) %>%
  left_join(totalkcal) %>%
  group_by(UserName, RecallNo, klass) %>%
  mutate(klass_ppintake = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, klass, klass_ppintake, avg_total_kcal)) %>%
  ungroup() %>%
  group_by(UserName, klass) %>% #Averages across recalls for each food
  mutate(Avg_klass_ppintake = mean(klass_ppintake/(avg_total_kcal/1000), na.rm = TRUE)) %>%
  distinct(UserName, klass, .keep_all = TRUE) %>%
  select(-c(RecallNo, klass_ppintake)) %>%
  group_by(klass) %>%
  add_count(klass) %>% #how many users per class
  mutate(overall_avg_klass_ppintake = mean(Avg_klass_ppintake, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
#how many classes did participants eat?
participantclass = class_analysis %>%
  group_by(UserName) %>%
  add_count(UserName, name = 'classes_consumed') %>%
  select(UserName, classes_consumed) %>%
  distinct(UserName, .keep_all = TRUE)

message("Mean Number of classes consumed: ", 
        round(mean(participantclass$classes_consumed), digits =0),
        "\nSD of classes consumed:", 
        round(sd(participantclass$classes_consumed), digits =0))
```

What are the top class contributors that have non-zero content in FooDB?
```{r}
summarizedclass = class_analysis %>%
  select(c(klass, overall_avg_klass_ppintake, n)) %>%
  distinct(klass, .keep_all = TRUE) %>%
  rename('klass_freq' = 'n') %>%
  filter(!overall_avg_klass_ppintake == 0) %>%
  arrange(desc(overall_avg_klass_ppintake))

head(summarizedclass$klass, 10)
```



