---
title: "FIG_Participant_Polyphenol_Summary"
author: "Stephanie Wilson"
date: "September 2023"
output: html_document
---

# Summarized Polyphenol Figure by Covariate Bins

This script looks at intakes of total polyphenols and polyphenolic classes and provides a summary across age, BMI, and sex bins. This summary is plotted as a stacked bar plot. 

__Inputs__. 
- *PP_Raw* - polyphenol content data linked to dietary records, Output from A0B_Exploration.Rmd.
- *Polyphenol taxonomy* - taxonomic classifications of polyphenols. Output from A0A_Scrape_Taxonomy.Rmd.
- *merged variables* - metadata from the FL100 study cohort. Output from A0_Merge_Variables.Rmd
  
__Outputs__. 
- *Stacked Bar Plot* - Polyphenol intake summaries With Sample sizes across covariate bins
  

Load packages
```{r message = FALSE}
library(tidyverse); library(Rfit); library(Hmisc); library(corrplot); library(ggpubr);
library(ggrain); library(viridis); library(ggridges); library(gt); library(wesanderson)
```

Load Functions
```{r}
# Summary stats by group
source('Functions/calculate_summary_stats.R')

# Function for summarizing polyphenol classes across variables of interest.
# Assumes data input has class polyphenol data and variables of interest
summarize_class_intake = function(data, group_column) {
  result = data %>%
    select(c(klass_label, Avg_klass_ppintake, {{group_column}})) %>%
    
    # Summarize across all participants
    group_by({{group_column}}, klass_label) %>%
    mutate(overall_avg_klass_ppintake = mean(Avg_klass_ppintake, na.rm = TRUE)) %>%
    ungroup() %>%
    distinct({{group_column}}, klass_label, .keep_all = TRUE) %>%
    select(-Avg_klass_ppintake) %>%
    
    # Summarize across group_column bins
    group_by({{group_column}}) %>%
    mutate(proportion = overall_avg_klass_ppintake / sum(overall_avg_klass_ppintake)) %>%
    ungroup() %>%
    
    # Filter out zeros
    filter(!overall_avg_klass_ppintake == 0) %>%
    arrange(desc(overall_avg_klass_ppintake))
  
  return(result)
}
```


```{r Load Data}
#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list)

# Load the Binning Data
bindata = read.csv('data/FL100_merged_variables.csv') %>%
  select(c(UserName, Sex, Age.Category, BMI.Category, avg_total_kcal)) %>%
  rename('BMI' = 'BMI.Category',
         'Age' = 'Age.Category') %>%
  
  #Filter in those who had quality controlled recalls
  filter(UserName %in% PP_Raw$UserName) %>%
  
  #Recode variables for clarity
  mutate(Age = recode_factor(Age,
                             "1" = "18 - 33.99",
                             "2" = "34 - 49.00",
                             "3" = "50 - 65"),
         BMI = recode_factor(BMI,
                             "1" = "18.5 - 24.99",
                             "2" = "25 - 29.99",
                             "3" = "30 - 39.99"))

# Polyphenol Taxonomy
tax = read.csv('FooDB/FooDB_phenols_taxonomy.csv')
```

### 1) Bin Size
```{r}
table(bindata$Sex)
table(bindata$BMI)
table(bindata$Age)
```

### 2) Total Polyphenol Breakdown

Obtain total intakes
```{r}
total_daily_intake = PP_Raw %>%
  group_by(UserName, RecallNo) %>%
  mutate(total_daily_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, total_daily_intake_mg)) %>%
  ungroup() %>%
  group_by(UserName) %>%
  mutate(Avg_Total_Daily_PP_Intake_mg = mean(total_daily_intake_mg, na.rm = TRUE)) %>%
  distinct(UserName, .keep_all = TRUE) %>%
  select(-c(RecallNo, total_daily_intake_mg)) %>%
  left_join(bindata, by = 'UserName') %>%
  arrange(UserName)
```

Age Summary
```{r}
calculate_summary_stats(data = total_daily_intake, group_var = Age, summary_var = Avg_Total_Daily_PP_Intake_mg)
```

BMI Summary
```{r}
calculate_summary_stats(data = total_daily_intake, group_var = BMI, summary_var = Avg_Total_Daily_PP_Intake_mg)
```

Sex Summary
```{r}
calculate_summary_stats(data = total_daily_intake, group_var = Sex, summary_var = Avg_Total_Daily_PP_Intake_mg)
```


### 3) CLASS SUMMARY
```{r}
# Average class consumed for each participant
class_analysis = PP_Raw %>%
  left_join(tax) %>%
  left_join(bindata) %>%
  
  #Average per recall
  group_by(UserName, RecallNo, klass) %>%
  mutate(klass_ppintake = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, Age, BMI, Sex, RecallNo, klass, 
           klass_ppintake, avg_total_kcal)) %>%
  ungroup() %>%
  
  #Averages across recalls for each food
  group_by(UserName, klass) %>%
  mutate(Avg_klass_ppintake = mean(klass_ppintake/(avg_total_kcal/1000), 
                                   na.rm = TRUE)) %>%
  distinct(UserName, klass, .keep_all = TRUE) %>%
  select(-c(RecallNo, klass_ppintake)) %>%
  group_by(klass) %>%
  add_count(klass) %>% #how many users per class
  mutate(overall_avg_klass_ppintake = mean(Avg_klass_ppintake, na.rm = TRUE)) %>%
  ungroup() %>%
  
  # Add new label for plotting
  mutate(klass_label = ifelse(klass %in% c('Flavonoids', 'Organooxygen compounds', 
                                           'Benzene and substituted derivatives', 
                                           '2-arylbenzofuran flavonoids',
                                           'Cinnamic acids and derivatives',
                                           'Alcohols and polyols', 'Isoflavonoids', 
                                           'Tannins', 'Phenols'), klass, 'Other'))
```



What is listed as other? 
 - others_list: Must be non-zero
 - others_zero_list: Must be zero
```{r}
others_list = class_analysis %>% filter(klass_label == 'Other' & overall_avg_klass_ppintake> 0) %>% select(c(klass_label, klass)) %>% distinct(klass, .keep_all = TRUE) %>% arrange(klass) %>% mutate(klass = str_to_lower(klass))

cat('Non-Zero Classes\n', paste(others_list$klass, collapse = ', '))

others_zero_list = class_analysis %>% filter(klass_label == 'Other' & overall_avg_klass_ppintake == 0) %>% select(c(klass_label, klass)) %>% distinct(klass, .keep_all = TRUE) %>% arrange(klass) %>% mutate(klass = str_to_lower(klass))

cat('\n\nZeroed Classes\n', paste(others_zero_list$klass, collapse = ', '))
```



How many classes did participants eat?
```{r}
participantclass = class_analysis %>%
  group_by(UserName) %>%
  add_count(UserName, name = 'classes_consumed') %>%
  select(UserName, classes_consumed) %>%
  distinct(UserName, .keep_all = TRUE)
```



### 4) Prepare Dataframes

Total
```{r}
#classes across all participants
summary_total = class_analysis %>%
  select(c(klass_label, overall_avg_klass_ppintake)) %>%
  distinct(klass_label, .keep_all = TRUE) %>%
  arrange(desc(overall_avg_klass_ppintake)) %>%
  filter(!overall_avg_klass_ppintake == 0) %>%
  mutate(proportion = overall_avg_klass_ppintake/sum(overall_avg_klass_ppintake),
         axis_label = 'All') %>%
  arrange(desc(proportion))
```

```{r}
message('Energy-Adjusted Intakes\nMean: ', mean(summary_total$Avg_klass_ppintake/(class_analysis$avg_total_kcal/1000)),
        '\nStandard Error: ', sd(class_analysis$Avg_klass_ppintake)/sqrt(nrow(class_analysis)))
```

Age, BMI, Sex
```{r}
#classes intake across different covariates
summary_age = summarize_class_intake(data = class_analysis, group_column = Age) %>%
  mutate(axis_label = paste0('Age: ', Age)) %>%
  select(-Age)

summary_BMI = summarize_class_intake(data = class_analysis, group_column = BMI) %>%
  mutate(axis_label = paste0('BMI: ', BMI)) %>%
  select(-BMI)

summary_Sex = summarize_class_intake(data = class_analysis, group_column = Sex) %>%
  mutate(axis_label = paste0('Sex: ', Sex)) %>%
  select(-Sex)
```

Combine all dataframes for a singular plot
```{r}
full_summary = full_join(summary_total, summary_age) %>%
  full_join(summary_BMI) %>%
  full_join(summary_Sex)
```


### 5) PLOTS

Labels
```{r}
# Find sum for each level for labelling at the top of the bar chart
total_labels = full_summary %>%
  group_by(axis_label) %>%
  mutate(total = round(sum(overall_avg_klass_ppintake), digits =0)) %>%
  ungroup() %>%
  distinct(axis_label, .keep_all = TRUE) %>%
  mutate(size = c('(n = 350)', '(n = 112)', '(n = 121)', '(n = 117)', 
                            '(n = 125)', '(n = 135)', '(n = 90)', '(n = 186)',
                            '(n = 164)')) %>%
  mutate(axis_label2 = paste(axis_label, size, sep = '\n'))

# Update the x-axis tick labels to include sample size
full_summary2 = total_labels %>%
  select(axis_label, axis_label2) %>%
  right_join(full_summary)
```
Colors
```{r}
# Select 10 colors from the Asteroid City 3 color palette
wes_colors = wes_palette("AsteroidCity3", type = 'continuous', 10)
```


```{r}
summary_plot = ggplot(full_summary2, aes(x = axis_label2, y =  overall_avg_klass_ppintake, fill = klass_label)) +
  geom_bar(stat = "identity", position = position_stack()) +
  geom_text(data = total_labels, aes(x = axis_label2, y = total, label = total, 
                                     hjust = 'middle'), nudge_y = 70, size =3) +
  labs(x = "", y = "Average Intake (mg/1000 kcal)", title = "") +
  theme_classic() +
  labs(fill = 'Polyphenol Class') +
  scale_fill_manual(values = wes_colors)+
  coord_flip() +
  scale_x_discrete(limits = c("Sex: Male\n(n = 164)", "Sex: Female\n(n = 186)", 
                              "BMI: 30 - 39.99\n(n = 90)", "BMI: 25 - 29.99\n(n = 125)",
                              "BMI: 18.5 - 24.99\n(n = 135)", "Age: 50 - 65\n(n = 112)",
                              "Age: 34 - 49.00\n(n = 117)", "Age: 18 - 33.99\n(n = 121)",
                              "All\n(n = 350)")) + 
  theme(axis.title.x = element_text(face = "bold", size = 12),
         legend.title = element_text(face = "bold"))

summary_plot
```

Save plots
```{r}
ggsave(plot = summary_plot, filename = 'images/polyphenol_summary_stackedbar.png', 
        width = 7, height = 4, units = 'in', dpi = 300)
ggsave(plot = summary_plot, filename = 'images/polyphenol_summary_stackedbar.pdf', 
        width = 7, height = 4, units = 'in')
```

