---
title: "Total Polyphenol and Inflammation Analysis"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Total Polyphenol Intake Analysis with Inflammation Markers

__Information__. 
This script compiles and plots model summaries from final models from A2 Blood and and A3 Fecal Analyses (which feature energy-adjusted total polyphenolic intake as a predictor).

__Required Input Files__.                          
  - *FL100_PP_Raw.csv* - Polyphenol content for each compound in foods listed in each recall test. Output from A0B_Exploration.Rmd.
  - *FL100_merged_variables.csv* - This merged file includes binning data, SES, dietary, blood, and stool data. Output from A0_Merge_Variables.Rmd

__Output__
  - RDS - Heatmap summary
  - Table Summary (Supplementary Table)

```{r Load packAges, message = FALSE, warning = FALSE}
library(tidyverse); library(Rfit); library(Hmisc); library(corrplot); library(ggpubr)
library(psych); library(car)
source('Functions/summarizemodels.R')
```


```{r Load Data}
#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

# Load in Dietary Polyphenol Content
# Raw is Each compound for each food in each recall by participant
PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list)

#multiple files including ASA averAges for kcal and fiber
merged = read.csv('data/FL100_merged_variables.csv') %>%
  select(c(UserName, Age, Sex, BMI, avg_total_kcal, avg_total_fiber,
           hei_asa24_totalscore, plasma_lbp_bd1, CRP_BD1, 
           After24h, fecal_neopterin, fecal_calprotectin, fecal_mpo)) %>%
  mutate(Sex = factor(Sex))
```



### 1) Merge Data

Sum polyphenol intake (mg) consumed for each recall, for each person, then averAged across recalls.
```{r}
#total_daily_intake = total daily intake (mg) for each participant's recall
total_daily_intake = PP_Raw %>%
  group_by(UserName, RecallNo) %>%
  mutate(total_daily_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>%
  select(c(UserName, RecallNo, total_daily_intake_mg)) %>%
  ungroup() %>%
  group_by(UserName) %>%
  mutate(Avg_Total_Daily_PP_Intake_mg = mean(total_daily_intake_mg)) %>%
  distinct(UserName, .keep_all = TRUE) %>%
  select(-c(RecallNo, total_daily_intake_mg)) %>%
  left_join(merged, by = 'UserName') %>%
  arrange(UserName) %>%
  relocate(Sex, After24h, .after = UserName) %>%
  mutate(energy_adj_Avg_Total_Daily_PP = Avg_Total_Daily_PP_Intake_mg/
           (avg_total_kcal/1000))
```

Look for missing variables
```{r}
total_daily_intake_CRP = total_daily_intake %>%
  filter(!is.na(CRP_BD1))
```


### 2) BLOOD MARKERS

#### CRP

CRP without Dietary Predictors
```{r}
#Removal of UserName 8048, row 256, outlier
CRP_model2 = rfit(CRP_BD1 ~ energy_adj_Avg_Total_Daily_PP + Age + BMI + Sex, data = total_daily_intake_CRP %>% filter(!UserName == 8048))
```

CRP WITH Dietary Predictors
```{r}
CRP_model_diet2 = rfit(CRP_BD1 ~ energy_adj_Avg_Total_Daily_PP + Age + Sex + BMI + avg_total_fiber + hei_asa24_totalscore, data = total_daily_intake_CRP %>% filter(!UserName==8048))
```

#### LBP
LBP without Dietary Predictors
```{r}
LBP_model = rfit(plasma_lbp_bd1 ~ energy_adj_Avg_Total_Daily_PP + Age + BMI + Sex, data = total_daily_intake)
```

LBP WITH Dietary Predictors
```{r}
LBP_model_diet = rfit(plasma_lbp_bd1 ~ energy_adj_Avg_Total_Daily_PP + Age + BMI + Sex + avg_total_fiber +hei_asa24_totalscore, data = total_daily_intake)
```

### 3) FECAL MARKERS

Filter for stool samples older than 24 hours. Look for missing variables
```{r}
total_daily_intake_cleaned = total_daily_intake %>% #350 starting
  filter(!After24h == 1) %>% #removes 30 individuals 
  filter(!is.na(fecal_neopterin)) #removes an additional 13
```

#### NEO
NEO without Dietary Predictors
```{r}
NEO_rfit = rfit(fecal_neopterin ~ energy_adj_Avg_Total_Daily_PP + Age + BMI + Sex, data = total_daily_intake_cleaned)
```

NEO WITH Dietary Predictors
```{r}
NEO_rfit_diet = rfit(fecal_neopterin ~ energy_adj_Avg_Total_Daily_PP + Age +
                 Sex + BMI + avg_total_fiber +
                 hei_asa24_totalscore, data = total_daily_intake_cleaned)
```


#### MPO

MPO without Dietary Predictors
```{r}
MPO_model = rfit(fecal_mpo ~energy_adj_Avg_Total_Daily_PP + Age + Sex + BMI, data = total_daily_intake_cleaned)
```

MPO WITH Dietary Predictors
```{r}
MPO_model_diet = rfit(fecal_mpo ~energy_adj_Avg_Total_Daily_PP + Age + Sex + BMI + avg_total_fiber + hei_asa24_totalscore, data = total_daily_intake_cleaned)
```

#### Calprotectin

CAL without Dietary Predictors
```{r}
CAL_rfit = rfit(fecal_calprotectin ~ energy_adj_Avg_Total_Daily_PP + Age + 
                 Sex + BMI, data = total_daily_intake_cleaned)
summary(CAL_rfit)

 plot(fitted(CAL_rfit), rstudent(CAL_rfit))
```

CAL WITH Dietary Predictors
```{r}
CAL_rfit_diet = rfit(fecal_calprotectin ~ energy_adj_Avg_Total_Daily_PP + 
                      Age + Sex   + BMI + avg_total_fiber +
                 hei_asa24_totalscore, data = total_daily_intake_cleaned)
```


### 4) Summary Plots 
From all models with total polyphenol intake

```{r}
nodiet = list(CRP_model2, LBP_model, CAL_rfit, MPO_model, NEO_rfit)
  names(nodiet) = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO')
withdiet = list(CRP_model_diet2, LBP_model_diet, CAL_rfit_diet, MPO_model_diet, NEO_rfit_diet)
  names(withdiet) = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO')
```


Run the functions to aggregate the model summaries
```{r}
nodiet_summary = summarizemodels(nodiet, 
                                 model_names = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO'),
                                 diet_control = FALSE)

withdiet_summary = summarizemodels(withdiet, 
                                   model_names = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO'),
                                   diet_control = TRUE)

joint_summaries = rbind(nodiet_summary, withdiet_summary)
```


```{r}
joint_summaries_clean = joint_summaries %>%
  filter(!Variable == '(Intercept)') %>%
  
  # Nicely formatted Variable Names
  mutate(Variable = ifelse(Variable == 'SexMale', 'Sex, Male', 
                           ifelse(Variable == 'energy_adj_Avg_Total_Daily_PP', 
                                  'Total Polyphenol Intake',
                                  ifelse(Variable == 'avg_total_fiber', 'Total Fiber',
                                         ifelse(Variable == 'hei_asa24_totalscore', 
                                                'ASA24 HEI Total Score', Variable)))),
    Estimate = round(Estimate, 3),
    p.value = signif(p.value, 3),
         Significance =  ifelse(p.value < 0.05, 'TRUE', 'FALSE'),
    
    # If a covariate is important, tag it with a special character    
    MarkerSig = ifelse(Variable == 'Age' & Significance == TRUE, "\u2020",
                            ifelse(Variable =='BMI' & Significance == TRUE, "\u2021",
                            ifelse(Variable =='Sex, Male' & Significance == TRUE, '\u00B6',
                            ifelse(Variable == 'Total Fiber' & Significance == TRUE, '#',
                            ifelse(Variable == 'ASA24 HEI Total Score' & Significance == TRUE,
                                   '\u2660', ''))))),
    #Return CRP, LBP, and MPO to normal scale
    Estimate_update = ifelse(outcome %in% c('CRP', 'MPO', 'LBP'),
                             exp(Estimate), Estimate))


combinedsymbols = joint_summaries_clean %>%
  group_by(outcome, diet_control) %>%
  # Merge the covariate symbols into one for each model
  # This is for presentation of Beta coefficients
  # This highlights the other relevant covariates in the same model
  reframe(CombinedMarkerSig = paste(MarkerSig, collapse = '')) %>%
  ungroup() %>%
  right_join(joint_summaries_clean) %>%
  mutate(Label = ifelse(Significance ==TRUE, paste(Estimate, ' ', 
                                                   CombinedMarkerSig, sep = ''), CombinedMarkerSig),
         outcome = factor(outcome, levels = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO')),
         diet_control = ifelse(diet_control == 'With Dietary Predictors', 
                               "With Fiber, Total HEI", "Without Fiber, Total HEI"),
         Variable = ifelse(Variable=="Total Polyphenol Intake", 'Total Polyphenol', Variable))# Add a default case if needed
```


```{r}
ALL_Sig = combinedsymbols %>%
  filter(Variable == 'Total Polyphenol' & Significance ==TRUE)

heatmaps_total = ggplot(combinedsymbols %>%
                    filter(Variable == 'Total Polyphenol'), 
                  aes(x = 'Total Polyphenol', y = outcome, fill = Estimate)) + 
  geom_tile(aes(label = Estimate)) +
   facet_grid(~ diet_control) +
  geom_text(data = ALL_Sig, 
            aes(label = Label, fontface = ifelse(outcome == 'CAL' & 
                                                   diet_control == 'Without Dietary Predictors',
                                                 'bold', 'plain')), size = 3) + 
  scale_fill_gradient2(low = "#CB7A5C",
                       mid = "lightgray",
                       high = "#5785C1",
                       name = expression(paste(beta, "-Coefficient"))) + 
  scale_color_manual(values = c('black', 'white')) +
  theme_bw() +
  labs(x = '', y = '') +
  coord_flip() + 
  theme(legend.position = "bottom", 
        legend.text = element_text(angle = 45,vjust = 1, hjust = 1))
 

heatmaps_total
```


```{r}
saveRDS(heatmaps_total, file = "images/Heatmap_TotalPP.rds")

#Save as image files for publication
ggsave(heatmaps_total, filename = 'images/TotalPolyphenolModels.pdf', height = 3 , width = 8, units = 'in')
ggsave(heatmaps_total, filename = 'images/TotalPolyphenolModels.png', height = 3 , width = 8, units = 'in')
```

#### Now let's create a table.
```{r}
table_data = combinedsymbols %>%
  mutate(p.value = ifelse(p.value < 0.001, '< 0.001',
                     ifelse(p.value < 0.01, '< 0.01', round(p.value, digits = 2))),
                          table_label = ifelse(Significance == TRUE, paste(Variable, "\u03B2 =", 
                                        paste(Estimate, ',', sep =''),
                                        'P', p.value, sep = ' '), '')) %>%
  select(outcome, diet_control, Variable, Estimate, p.value, table_label) %>%
  mutate(diet_control = ifelse(diet_control == 'With Fiber, Total HEI', TRUE, FALSE)) 

#Compile covariate relationships
 covariate= table_data %>%
  filter(!Variable == 'Total Polyphenol') %>% 
   group_by(outcome, diet_control) %>%
  reframe('Covariate Relationships' = paste(table_label, collapse = "")) 
```


```{r}
#Merge covariate relationship with total polyphenol beta and pvalues
formatted_table_data = table_data %>%
   filter(Variable == 'Total Polyphenol') %>%
  select(-c(table_label, Variable)) %>%
  left_join(covariate) %>%
  rename("Total Polyphenol \u03B2" = 'Estimate',
         Marker = outcome,
         'Total Polyphenol p-value' = p.value) %>%
  mutate(Marker = factor(Marker, levels = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO'))) %>% 
  arrange(Marker, diet_control) %>%
  relocate(diet_control, .before = Marker)

write.csv(formatted_table_data, 'output/TotalPolyphenolModelsTABLE.csv', row.names =FALSE, 
          fileEncoding = "UTF-8")
```

