---
title: "A4_ANCOVAs_ClassTransformed"
author: "Stephanie Wilson"
date: "September 2023"
output: pdf_document
---

# Class Level Analyses with Inflammation Markers

__Inputs__
  - FL100_merged_variables.csv - Output from Other_Merged_Variables
  - taxonomy.csv - taxonomic classifications for polyphenolic compounds
  - FL100_PP_Raw.csv - polyphenolic data for each compound, recall, and participant
  
__Summary__
The following code runs repeated ANCOVA tests with multiple testing adjustment in regards to assessment of polyphenolic class and five outcomes (CRP, LBP, NEO, MPO, CAL). This is done both with and without dietary predictors (fiber intake and total HEI score). Each polyphenolic class is transformed by the BestNormalize recommendation. In essence, the code runs as follows in an example without dietary predictors:

  Outcome1 ~ transformation(class1) + Age + Sex + BMI
  Outcome1 ~ transformation(class2) + Age + Sex + BMI
  ...
  Outcome5 ~ transformation(class1) + Age + Sex + BMI
  
1) Best Normalize on Taxa
2) Run ANCOVAs for each Outcome
3) Merge Data Into One File

__Output__
- All_ANCOVA_results.csv
- FooDB_phenol_content_DailyClassAvg.csv

Load packages
```{r warning = FALSE, message = FALSE}
#Load packages
library(tidyverse); library(car); library(bestNormalize); library(stringr)

# ANCOVA with multiple testing adjustments for analyzing multiple taxa on outcomes
source('Functions/multipleANCOVA.R')
source('Functions/find_transformation.R')
```

Load Data
```{r}
taxonomy = read.csv('FooDB/FooDB_phenols_taxonomy.csv') %>%
  select(c(compound_public_id, kingdom:klass, compound_name)) %>%
  mutate(klass = ifelse(klass =='', paste('Unclassified', superklass), klass))

#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

# Load in Dietary Polyphenol Content, 
# Raw is Each compound for each food in each recall by participant
PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list)
```

### 1) Calculate Average Intake for every taxonomic class from each participant
```{r}
class =  PP_Raw %>%
  left_join(taxonomy, by = 'compound_public_id') %>%
  
  #Group by Taxonomic Class
  group_by(UserName, RecallNo, klass) %>%
  
  #gets the sum of each compound for each participant's recall
  mutate(klass_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>% 
  select(c(UserName, RecallNo, klass, klass_intake_mg)) %>%
  ungroup()%>%
  
  #Remove duplicates since we've summed each polyphenol per recall
  distinct(UserName, RecallNo, klass, .keep_all = TRUE) %>%
  
  #Average polyphenol intake across recalls for each compound
  group_by(UserName, klass) %>%
  mutate(Avg_klass_intake = mean(klass_intake_mg)) %>%
  ungroup() %>%
  
  #Remove duplicates
  distinct(UserName, klass, .keep_all = TRUE) %>%
  select(c(UserName, klass, Avg_klass_intake)) %>%
  filter(!is.na(klass)) %>%
  
  #Transpose dataframe where each column is a participant
  pivot_wider(id_cols = UserName, names_from = klass, 
              values_from = Avg_klass_intake, values_fill = 0) %>%
  
  #remove the columns that are just zeros
  select(-c('Unclassified':'Thiolanes'))
```


```{r}
write.csv(class, 'FooDB/FooDB_phenol_content_DailyClassAvg.csv', 
          row.names = FALSE)
```


Class Filtering, Remove Taxa with more than 80% as zeros.
```{r}
toremove = as.data.frame(colSums(class[ , -1] == 0)) %>%
  rename('zero_counts' = 1) %>%
  mutate(proportion = zero_counts/350) %>%
  rownames_to_column(var = 'Tax_class') %>%
  filter(proportion > 0.2) %>%
  select(Tax_class)

#Convert from dataframe.
remove = toremove$Tax_class

# Update class to include this filtering
updated_class = class %>%
  select(!all_of(toremove$Tax_class))
```


Load the metadata and merge with taxonomic data.
```{r}

# Multiple files including ASA averages for kcal and fiber
merged = read.csv('data/FL100_merged_variables.csv') %>%
  select(c(UserName, Age, Sex, BMI, avg_total_kcal, avg_total_fiber,
           hei_asa24_totalscore, After24h, CRP_BD1, plasma_lbp_bd1,
           fecal_calprotectin, fecal_mpo, fecal_neopterin)) %>%
  mutate(Sex = factor(Sex),
         After24h = factor(After24h)) %>%
  filter(!is.na(avg_total_kcal)) %>%
  left_join(updated_class, by = 'UserName')

merged_fecal = merged %>%
  filter(!After24h == 1) %>% #removes 30 individuals 
  filter(!is.na(fecal_neopterin))

merged_blood = merged %>%
  filter(!is.na(CRP_BD1))
```

  
### 1) Best Normalize on Taxa

```{r}
# Create a vector of taxa names
## Class Names are the same for both blood and fecal datasets
class_names = colnames(merged_blood)[14:28]
```


Find the recommended bestNormalize transformation for each of polyphenolic classes.
```{r}
blood_transformations = find_transformation(data = merged_blood, class_names = class_names)
fecal_transformations = find_transformation(data = merged_blood, class_names = class_names)

# If there are two transformations for Flavonoids
#blood_transformations = blood_transformations %>% filter(!Class == 'Flavonoids' | !transformation == 'yeojohnson')
#blood_transformations = blood_transformations %>% filter(!Class == 'Flavonoids' & !transformation == 'yeojohnson')
```


### 2) ANCOVAs

#### CRP
```{r}
CRP = multipleANCOVA(data = merged_blood, class_names = class_names, outcome = merged_blood$CRP_BD1, diet_control = FALSE, 
                     transformations = blood_transformations)

CRPdiet = multipleANCOVA(data = merged_blood, class_names = class_names, outcome = merged_blood$CRP_BD1, diet_control = TRUE,
                         transformations = blood_transformations)

CRP$Outcome = 'CRP'
CRPdiet$Outcome = 'CRP'
```

#### LBP
```{r}
LBP = multipleANCOVA(data = merged_blood, class_names = class_names, outcome = merged_blood$plasma_lbp_bd1, diet_control = FALSE,
                      transformations = blood_transformations)

LBPdiet = multipleANCOVA(data = merged_blood, class_names = class_names, outcome = merged_blood$plasma_lbp_bd1, diet_control = TRUE,
                          transformations = blood_transformations)

LBP$Outcome = 'LBP'
LBPdiet$Outcome = 'LBP'
```

#### NEO
```{r}
NEO = multipleANCOVA(data = merged_fecal, class_names = class_names, outcome = merged_fecal$fecal_neopterin, diet_control = FALSE,
                     transformations = fecal_transformations)

NEOdiet = multipleANCOVA(data = merged_fecal, class_names = class_names, outcome = merged_fecal$fecal_neopterin, diet_control = TRUE,
                         transformations = fecal_transformations)

NEO$Outcome = 'NEO'
NEOdiet$Outcome = 'NEO'
```

#### CAL
```{r}
CAL = multipleANCOVA(data = merged_fecal, class_names = class_names, outcome = merged_fecal$fecal_calprotectin, diet_control = FALSE,
                     transformations = fecal_transformations)

CALdiet = multipleANCOVA(data = merged_fecal, class_names = class_names, outcome = merged_fecal$fecal_calprotectin, diet_control = TRUE,
                         transformations = fecal_transformations)

CAL$Outcome = 'CAL'
CALdiet$Outcome = 'CAL'
```

#### MPO
```{r}
MPO = multipleANCOVA(data = merged_fecal, class_names = class_names, outcome = merged_fecal$fecal_mpo, diet_control = FALSE,
                     transformations = fecal_transformations)

MPOdiet = multipleANCOVA(data = merged_fecal, class_names = class_names, outcome = merged_fecal$fecal_mpo, diet_control = TRUE,
                         transformations = fecal_transformations)

MPO$Outcome = 'MPO'
MPOdiet$Outcome = 'MPO'
```


### 3) Merge Data

```{r message = FALSE}
NoDiet = CRP %>%
  full_join(LBP) %>%
  full_join(NEO) %>%
  full_join(CAL) %>%
  full_join(MPO) %>%
  relocate(Outcome, .before = Class) %>%
  mutate(diet_control = "No")

WithDiet = CRPdiet %>%
  full_join(LBPdiet) %>%
  full_join(NEOdiet) %>%
  full_join(CALdiet) %>%
  full_join(MPOdiet) %>%
  relocate(Outcome, .before = Class) %>%
  mutate(diet_control = "Yes")
```


Combine both datasets and Format for Cleaner Presentation
```{r}
All_ANCOVA_results = full_join(NoDiet, WithDiet) %>%
  relocate(diet_control, .after = "Outcome") %>% 
  arrange(Adjusted_P_Value) %>%
  mutate(diet_control = case_when(
      diet_control == 'Yes' ~ "With Dietary Predictors",
      diet_control == 'No' ~ "Without Dietary Predictors",
      TRUE ~ diet_control),
      Class = str_replace_all(Class, "\\.", " "),
      Class = ifelse(Class == "X2 arylbenzofuran flavonoids", 
                     '2 arylbenzofuran flavonoids', Class),
      Class_transformation = paste("(", Class_transformation, ")", sep = "")) %>%
      unite('Class_label',Class:Class_transformation, sep = ' ',remove = FALSE) %>%
  #filter(!Variable == '(Intercept)') %>%
  
  # Nicely formatted Variable Names
  mutate(Variable = ifelse(Variable == 'SexMale', 'Sex, Male', 
                    ifelse(Variable == 'class_transform', 'Polyphenol Class Intake',
                    ifelse(Variable == 'avg_total_fiber', 'Total Fiber',
                    ifelse(Variable == 'hei_asa24_totalscore', 'ASA24 HEI Total Score',
                           Variable)))),
         Significance =  ifelse(Adjusted_P_Value < 0.05, 'TRUE', 'FALSE'),
    
    # If a covariate is important, tag it with a special character    
    MarkerSig = ifelse(Variable == 'Age' & Significance == TRUE, "\u2020",
                ifelse(Variable =='BMI' & Significance == TRUE, "*",
                ifelse(Variable =='Sex, Male' & Significance == TRUE, '\u00B6',
                ifelse(Variable == 'Total Fiber' & Significance == TRUE, '#',
                ifelse(Variable == 'ASA24 HEI Total Score' & Significance == TRUE,
                                   '\u2660', ''))))))

write.csv(All_ANCOVA_results, 'output/All_ANCOVA_results.csv', row.names = FALSE)
```

