---
title: "FIG_ANCOVA_Heatmap"
author: "Stephanie Wilson"
date: "September 2023"
output: html_document
---

__Information__. 
This script pulls in the outputs from the ANCOVA tests (where polyphenol classes are predictors, inflammation markers are a response) and creates a heatmap summary. 
*Requires* ANCOVAs_ClassTransformed.Rmd has been run. 

__Input Files__.                          
  - *All_ANCOVA_results.csv* - Polyphenol content for each compound in foods listed in each recall test. Output from A4_ANCOVAs_ClassTransformed.Rmd

__Output__
  - RDS Heatmap summary
  
```{r}
library(tidyverse); library(doBy)
```

Load Data
```{r}
All_ANCOVA_results = read.csv('output/All_ANCOVA_results.csv') %>%
  filter(!Variable =='(Intercept)') %>%
  mutate(diet_control = case_when(
           diet_control == 'With Dietary Predictors' ~ 'With Fiber, Total HEI',
           diet_control == 'Without Dietary Predictors' ~ 'Without Fiber, Total HEI'),
         location = case_when(
           Outcome %in% c("MPO", "CAL", "NEO") ~ "fecal",
           Outcome %in% c("CRP", "LBP") ~ "plasma",
           TRUE ~ "other"  # You can set a default value for other cases if needed
         ),
         #Relevel markers
         Outcome = factor(Outcome, levels = c("CRP", "LBP", "CAL", "MPO", "NEO"))) 
```


### 1) Prepare combined 'significance' symbols for visual. 

Variable of interest is for polyphenol class.
```{r}
combinedsymbols = All_ANCOVA_results %>%
  group_by(Outcome, diet_control, Class) %>%
  # Merge the covariate symbols into one for each model
  # This is for presentation of Beta coefficients
  # This highlights the other relevant covariates in the same model
  reframe(CombinedMarkerSig = paste(MarkerSig, collapse = '')) %>%
  ungroup() %>%
  right_join(All_ANCOVA_results)  # Add a default case if needed
```
Calculate means of the absolute value of each Coefficient. Normalize polyphenol intake class coefficients to the calculated group mean.
```{r}
means = combinedsymbols %>% 
  filter(Variable=="Polyphenol Class Intake") %>%
  group_by(Outcome) %>%
  summarize(mean_coefficient = mean(abs(Coefficient))) %>%
  mutate(Variable = "Polyphenol Class Intake")

combinedsymbols_updated = combinedsymbols %>%
  left_join(means) %>%
  mutate(Coefficient_norm = ifelse(is.na(mean_coefficient), NA, Coefficient/mean_coefficient))
```


### 2) Heat Map

```{r}
# Call in those formatted labels for our important polyphenolic classes
# These will be used as labels for our heatmap tiles.
ALL_Sig = combinedsymbols_updated %>%
  filter(Variable == 'Polyphenol Class Intake' & Adjusted_P_Value < 0.05) %>%
  mutate(Label = paste(signif(Coefficient_norm, 2), ' ', CombinedMarkerSig, sep = '')) # Add a default case if needed

heatmaps = ggplot(combinedsymbols_updated %>% filter(Variable == 'Polyphenol Class Intake'), 
                  aes(x = Class, y = Outcome, fill = Coefficient_norm)) + 
  geom_tile(aes(label = Coefficient)) +
  geom_text(data = ALL_Sig, aes(label = Label), color = 'black', size = 3) + 
  facet_grid(~ diet_control) +
  scale_fill_gradient2(low = "#CB7A5C",
                       mid = "lightgray",
                       high = "#5785C1",
                       name = expression(atop(paste(beta, "-Coefficient"), "(mean-centered)"))) + 
  theme_bw()  + 
  coord_flip() + 
  #Move legend to bottom
  theme(legend.direction = "vertical", legend.position = "right") +
  labs(y = '', x ='Class')
 
heatmaps
```

Export the files
```{r}
# As a ggplot object
saveRDS(heatmaps, file = "images/Heatmap_ClassPP.rds")

#For publications
ggsave(heatmaps, filename = 'images/Outcome_ANCOVA.pdf', height = 6 , width = 8, units = 'in')
ggsave(heatmaps, filename = 'images/Outcome_ANCOVA.jpg', height = 6 , width = 8, units = 'in')
```

