---
title: "PUBFIG_Total_Class_RegressionPlots"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Create heatmap and regression plots for class analyses

__Required Input Files__.                          
  - *FL100_PP_Raw.csv* - Polyphenol content for each compound in foods listed in each recall test. Output from A0_FilePrep_Exploration.Rmd.
  - *FooDB_phenols_taxonomy.csv* - Output from A0A_Scrape_Taxonomy.Rmd
  - *FooDB_phenol_content_DailyClassAvg.csv* - Output from A4_Class_Analyses.Rmd, Average Intake for every compound from each Participant
  - *FL100_merged_variables.csv* - Compilation of FL100 variables from A0_Merge_Variables.Rmd
  - *All_ANCOVA_results.csv* - Output from A3_Class_Analyses.Rmd
  
__Outputs__
  - Class Polyphenol Analysis Plots, food sources
  
```{r Load packages, message = FALSE}
library(tidyverse); library(ggpubr); library(bestNormalize); library(smplot2)
```


```{r Load Data}
# Polyphenol compound classification
tax = read.csv('FooDB/FooDB_phenols_taxonomy.csv') %>%
  select(c(compound_public_id, klass))

#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list) %>%
  left_join(tax)

#Intakes by class
class = read.csv('FooDB/FooDB_phenol_content_DailyClassAvg.csv')

#multiple files including ASA averages for kcal and fiber
merged = read.csv('data/FL100_merged_variables.csv') %>%
  mutate(Sex = factor(Sex))
```

### 1) Prepare Data

Yeo-Johnson Transformation
```{r}
class$Prenol_yeo = bestNormalize(class$Prenol.lipids)$x.t
class$Phenylprop_yeo = bestNormalize(class$Phenylpropanoic.acids)$x.t
```

Class Data Load and Clean
```{r}
ANCOVA = read.csv('output/All_ANCOVA_results.csv') %>%
  filter(Outcome=='LBP' & Class %in% c('Prenol lipids', 'Phenylpropanoic acids')) %>%
  filter(Variable %in% c('(Intercept)', 'Polyphenol Class Intake')) 

#Isolate Intercepts
Intercepts = ANCOVA %>%
  filter(Variable == ('(Intercept)')) %>%
  select(c(diet_control, Class, Coefficient)) %>%
  rename('Intercept' =3 )

ANCOVA_filter = ANCOVA %>%
  filter(!Variable == ('(Intercept)')) %>%
  left_join(Intercepts)
```

Isolate food sources of prenol lipids and phenylpropanoic acids to generate intake profiles
```{r}
consumers = PP_Raw %>%
  # Filter in sunflower, olive, and nectarines
  filter(food_id %in% c(86, 229, 121)) %>%
  # Filter in phenylproponoic acids and prenol lipids
  filter(klass %in% c("Phenylpropanoic acids", "Prenol lipids")) %>%
  # SUM INTAKE FOR EACH POLYPHENOL CLASS BY FOOD, RECALL, USER
  group_by(UserName, RecallNo, food_id, klass) %>%
  mutate(SUM_class_intake = sum(pp_consumed, na.rm =TRUE)) %>%
  distinct(UserName, RecallNo, food_id, klass, .keep_all = TRUE) %>%
  ungroup() %>%
  select(UserName, RecallNo, food_id, food_name, klass, SUM_class_intake) %>%
  #Averages across recalls for each food, klass combo
  group_by(UserName, food_id, klass) %>% 
  mutate(AVG_class_intake = mean(SUM_class_intake, na.rm = TRUE)) %>%
  distinct(UserName, food_id, klass, .keep_all = TRUE) %>%
  ungroup() %>%
  select(-c(RecallNo, SUM_class_intake))

seed_nectarine_consumers = consumers %>%
  filter(food_id %in% c(86, 229)) %>%
  distinct(UserName) %>%
  mutate(consumer = 'yes')
```

### 2) Class Intake Plot
```{r}
intake_plot = ggplot(consumers, aes(x = food_name, y = AVG_class_intake)) +
  geom_boxplot(notch = TRUE)+
  geom_jitter(position=position_jitter(0.2), alpha = 0.2) +
  facet_wrap(~klass, scales = 'free_y') +
  labs(x = "",
       y = "Average Intake (g)",
       fill = "Food") +
  theme_classic() +
  theme(legend.position = "none",
        strip.background = element_rect(fill = "grey90"),
         text = element_text(size = 12),
        axis.text.x = element_text(angle = 45,vjust = 1, hjust = 1)) + 
  scale_color_grey()
```


### 3) Class Regression Figure
Filter in Prenol lipids and phenylpropanoic acids. Format to long format for plotting, clean up class names.
```{r}
sigclass = merged %>%
  filter(!is.na(plasma_lbp_bd1))%>%
  left_join(class, by = "UserName") %>%
  select(c(UserName, avg_total_kcal, plasma_lbp_bd1, 
           Prenol_yeo, Phenylprop_yeo)) %>%
  pivot_longer(cols =4:5, names_to = 'class', values_to = 'value') %>% 
  mutate(class = case_when(
   class == 'Prenol_yeo' ~ 'Prenol Lipids',
    class == 'Phenylprop_yeo' ~ 'Phenylpropanoic Acids')) %>%
  left_join(seed_nectarine_consumers, by = "UserName")
```

```{r}
# Feature the consumers of seeds and nectarines
consumer_points = sigclass %>%
  filter(consumer == "yes")

#Class plots
class_plots = ggscatter(sigclass_update, x = "value", y = "plasma_lbp_bd1", 
                        add = "reg.line") +
  facet_wrap(~str_wrap(class, width =25), scales = 'free_x', ncol=2) +
  geom_smooth(method = "lm", se = FALSE) +
  coord_cartesian(ylim = c(min(sigclass$plasma_lbp_bd1),
                           max(sigclass$plasma_lbp_bd1)*1.2)) +
   stat_cor(method = 'spearman', p.accuracy = 0.001, size =3,
           label.x.npc = 'left',
           label.y = max(sigclass$plasma_lbp_bd1)*1.1) +
  geom_point(data = consumer_points, 
             aes(x = value, y = plasma_lbp_bd1),
                 colour = "#E54E21") +
  labs(y = expression("LBP (µg/mL)"), 
       x = 'Intake (mg/d, Yeo-Johnson transformed)',
       colour = '') + 
  scale_colour_discrete(name = '', 
                      labels = "Consumed sunflower\nseeds, nectarines") +
  theme_classic() +
  theme(plot.margin = unit(c(0,0,0,0.5), "cm"),
        legend.position = "right", 
        strip.background = element_rect(fill = "grey90"),
        text = element_text(size = 12)) 
```


### 3) Produce Publication Plots
```{r}
ggsave(intake_plot, filename = 'images/Class_Food_Sources.pdf', height = 3, width = 4, units = 'in')
ggsave(intake_plot, filename = 'images/Class_Food_Sources.jpg', height = 3, width = 4, units = 'in', dpi = 300) 

ggsave(class_plots, filename = 'images/Class_Regression_Plots.pdf', height = 3, width = 5, units = 'in')
ggsave(class_plots, filename = 'images/Class_Regression_plots.jpg', height = 3, width = 5, units = 'in', dpi = 300) 
```

