---
title: "Compound HFE File Preparation"
author: "Stephanie Wilson"
date: "`r Sys.Date()`"
output: html_document
---

# File Preparation for HFE of Polyphenolic compounds 

This script prepares input and metadata files for use in the HFE program built by Dr. Andrew Oliver.

__Inputs__. 
  - *FL100_merged_variables.csv* - many collective files to reduce the amount of input files
  - taxonomy - From A0A_Scrape_Taxonomy.Rmd, taxonomic classifications for polyphenolic compounds
  - *FL100_PP_Raw.csv* - From A0B_Exploration.Rmd, polyphenolic data for each compound, recall, and participant
  
__Outputs__. 
  - metadata_file_fecal.csv (sample size adjusted)
  - metadata_file_blood.csv (sample size adjusted)
  - covariates.csv
  - input_file_raw.csv - Average intake for each polyphenol 
  - input_file_relabund.csv (KEY) - Average intake for each polyphenol, converted to relative abundance 
  
  
```{r LOAD PACKAGES, messages = FALSE}
library(dplyr); library(tidyr); library(purrr); library(bestNormalize)
library(rvest); library(xml2); library(stringr); library(Hmisc)
```

Load Data
```{r}
taxonomy = read.csv('FooDB/FooDB_phenols_taxonomy.csv') %>%
  select(c(compound_public_id, kingdom:klass, compound_name)) %>%
  mutate(klass = ifelse(klass =='', paste('Unclassified', superklass), klass))

#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

# Load in Dietary Polyphenol Content, 
# Raw is Each compound for each food in each recall by participant
PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list)

#multiple files including ASA averages for kcal and fiber
merged = read.csv('data/FL100_merged_variables.csv') %>%
  mutate(Sex = factor(Sex)) %>%
  rename('subject_id' = 'UserName')
```


### 1) Obtain Average Intake for every compound from each Participant
```{r}
total_daily_intake = PP_Raw %>%
  group_by(UserName, RecallNo, compound_public_id) %>%
  
  #gets the sum of each compound for each participant's recall
  mutate(compound_intake_mg = sum(pp_consumed, na.rm = TRUE)) %>% 
  select(c(UserName, RecallNo, compound_public_id, compound_intake_mg)) %>%
  ungroup()%>%
  
  #Remove duplicates since we've summed each polyphenol per recall
  distinct(UserName, RecallNo, compound_public_id, .keep_all = TRUE) %>%
  group_by(UserName, compound_public_id) %>%
  
  #Average polyphenol intake across recalls for each compound
  mutate(Avg_compound_intake = mean(compound_intake_mg, na.rm = TRUE)) %>%
  ungroup() %>%
  
  #Remove duplicates
  distinct(UserName, compound_public_id, .keep_all = TRUE) %>%
  select(c(UserName, compound_public_id, Avg_compound_intake))

#Let's go ahead and write this file
write.csv(total_daily_intake, 'FooDB/FooDB_phenol_content_DailyCompoundAvg.csv', row.names = FALSE)
```

### 2) Prepare Taxonomy Clade Name

```{r}
#Define a mapping of Greek letters and their replacements
greek_mapping = c("α" = "alpha", "β" = "beta", "γ" = "gamma", "δ" = "delta",
                   "ε" = "epsilon", "ζ" = "zeta", "η" = "eta", "θ" = "theta",
                   "ι" = "iota", "κ" = "kappa", "λ" = "lambda", "μ" = "mu",
                   "ν" = "nu", "ξ" = "xi", "ο" = "omicron", "π" = "pi",
                   "ρ" = "rho", "σ" = "sigma", "τ" = "tau", "υ" = "upsilon",
                   "φ" = "phi", "χ" = "chi", "ψ" = "psi", "ω" = "omega")

# Function to replace Greek letters using the mapping
replace_greek = function(match) {
  greek_mapping[[match]]
}
```

```{r}
#Replace missing variables as Unclassified
taxonomy[taxonomy == ""] = 'Unclassified'   

#Prepare taxonomy labels and concatenate
taxonomy_clade = taxonomy %>%
  
  #CLEAN TAXONOMY STRINGS
  #mutate_if(is.character, str_replace_all, "[^[:alnum:]_]+", '_')  %>% #replace non alphanumeric except _
  #mutate_if(is.character, str_replace_all, "[α-ωΑ-Ω]", replace_greek) %>% #replace Greek letters
  #mutate_if(is.character, str_replace_all, '^_', '') %>% #remove leading _
  #mutate_if(is.character, str_replace_all, '_$', '') %>% #remove trailing _

  #ADD LABELS for Taxa HFE
  mutate(kingdom = sub("^", "k__", kingdom),
         superklass = sub("^", "p__", superklass),
         klass = sub("^", "c__", klass),
         compound_name = sub("^", "s__", compound_name)) %>%
  
  rename('species' = 'compound_name') %>%
  unite('clade_name', kingdom:species, sep = '|', remove = TRUE) #concatenate

write.csv(taxonomy_clade, 'HFE/input_file_clade_key.csv', row.names=FALSE)
```

### 3) Prepare the input dietary files for HFE

#### 3a) Input_file, RAW
```{r}
compound_intake = total_daily_intake %>%
  
  #Transpose dataframe where each column is a participant
  pivot_wider(names_from = UserName, values_from = Avg_compound_intake, 
              values_fill = 0) %>%
  left_join(taxonomy_clade, by = 'compound_public_id') %>%
  relocate('clade_name', .before = 'compound_public_id') %>%
  select(-compound_public_id) %>% 
  filter(!is.na(clade_name))

#write.table(compound_intake, 'HFE/input_file_raw.txt', sep = "\t", quote = FALSE, row.names = FALSE)
```

#### 3b) input_file, relative abundance
```{r}
# Calculate the sum of each participant's 'species' counts
participant_totals = colSums(compound_intake[-1])

# Divide each species count within a participant by the total count for that participant
relative_abundance = sweep(compound_intake[-1], 2, participant_totals, "/")

# Multiply the result by 100 to obtain relative abundance as a percentage
relative_abundance = relative_abundance * 100

# Add the clade_name back
compound_intake_relabund = relative_abundance %>%
  mutate(clade_name = compound_intake$clade_name) %>%
  relocate(clade_name, .before = 1)

write.table(compound_intake_relabund, 'HFE/input_file_relabund.txt', sep = "\t", quote = FALSE, row.names = FALSE)
```


#### 3c) metadata_file.txt for fecal measures

```{r}
fecal = merged %>%
  filter(!is.na(avg_total_kcal)) %>% #removes 43 from incomplete ASA24
  filter(!After24h == 1) %>% #removes 30 individuals 
  filter(!is.na(fecal_neopterin)) %>% #removes 13
  select(c(subject_id, fecal_calprotectin, fecal_mpo, fecal_neopterin)) %>%
  mutate(neo_quartile = as.numeric(cut2(fecal_neopterin, g = 4)),
         mpo_quartile = as.numeric(cut2(fecal_mpo, g = 4)),
         cal_quartile = as.numeric(cut2(fecal_calprotectin, g = 4))) 

#Add transformed variables
mpo_transformed = boxcox(fecal$fecal_mpo)
neo_transformed = boxcox(fecal$fecal_neopterin)

fecal = fecal %>%
  mutate(neo_boxcox = neo_transformed$x.t,
         mpo_boxcox = mpo_transformed$x.t)

write.table(fecal, 'HFE/metadata_fecal.txt', sep = "\t", quote = FALSE, row.names = FALSE)
```


#### 3c) metadata_file.txt for blood measures
```{r}
blood = merged %>%
  filter(!is.na(avg_total_kcal)) %>% #removes 43 from incomplete ASA24
  select(c(subject_id, CRP_BD1, plasma_lbp_bd1)) %>%
  filter(!is.na(CRP_BD1)) %>% # removes 2
  mutate(crp_quartile = as.numeric(cut2(CRP_BD1, g = 4)),
         lbp_quartile = as.numeric(cut2(plasma_lbp_bd1, g = 4)),
         crp_boxcox = boxcox(CRP_BD1)$x.t)
```


### SUBCLINICAL

Continuous
CRP < 10000 ng/mL
```{r}
metaCRP_subclin = blood %>% 
  filter(CRP_BD1 < 10000)

write.table(metaCRP_subclin, 'HFE/metadata_CRP_subclinical.txt',
            sep = "\t", quote = FALSE, row.names = FALSE)
```

Continuous - GI
Based off of fecal calprotectin < 100
```{r}
metafecal_subclin = fecal %>% 
  filter(fecal_calprotectin < 100) 

write.table(metafecal_subclin, 'HFE/metadata_fecal_subclinical.txt',
            sep = "\t", quote = FALSE, row.names = FALSE)
```



