---
title: "Pooled taxaHFE Features"
author: "Stephanie Wilson"
date: "October 2023"
output: html_document
---

# Aggregate and format names of taxaHFE-selected features

This script compiles the selected polyphenolic features from taxaHFE. 

__Inputs__. 
  - taxaHFE output.csv files
  - Taxonomy File, FooDB_phenols_taxonomy.csv
  
__Outputs__. 
  - HFE_pooled_selected_features_longform.csv
  - HFE_pooled_selected_features_compiled.csv

```{r message = FALSE}
library(tidyverse); library(stringdist)
source('functions/find_best_matches_taxonomy.R')
```


What were the input features for the Full Classifier?
```{r}
Inputs = data.frame(markers = c('CRP', 'LBP', 'CAL', 'MPO', 'NEO')) %>%
  mutate(path_name = paste('HFE/Subclinical/', markers, '_output', '.csv', 
                           sep ='')) %>%
  filter(!markers=='LBP')

taxonomy = taxonomy = read.csv('FooDB/FooDB_phenols_taxonomy.csv') %>%
  select(c(compound_public_id, kingdom:klass, compound_name)) %>%
  mutate(klass = ifelse(klass =='', paste('Unclassified', superklass), klass))
```

### 1) Read in each HFE output file, extract the column names, and compile into joint dataframe.
```{r}
# Create a list to store the dataframes
all_dataframes = data.frame()

# Loop through each file path
for (i in Inputs$path_name) {
  # Read the CSV file
  current_dataframe = read.csv(i, header = TRUE, sep = ",")

  # Check if the dataframe has any rows
  if (nrow(current_dataframe) > 0) {
    # Get the column names
    columns = names(current_dataframe)
    
    # Remove first two columns
    columns = columns[-c(1:2)]
    
    # Extract the filename without extension
    marker = gsub("\\.csv$", "", basename(i))
    marker = gsub('_quartile_output', '', marker)
    
    # Create a new dataframe with the extracted column names as rows
    columns_df = data.frame(features = columns, marker = marker)
    
    # Append the current dataframe to the list
    all_dataframes = rbind(columns_df, all_dataframes)
  }
}

```


### 2) Clean the Taxonomy Key

```{r}
#Define a mapping of Greek letters and their replacements
greek_mapping = c("α" = "alpha", "β" = "beta", "γ" = "gamma", "δ" = "delta",
                   "ε" = "epsilon", "ζ" = "zeta", "η" = "eta", "θ" = "theta",
                   "ι" = "iota", "κ" = "kappa", "λ" = "lambda", "μ" = "mu",
                   "ν" = "nu", "ξ" = "xi", "ο" = "omicron", "π" = "pi",
                   "ρ" = "rho", "σ" = "sigma", "τ" = "tau", "υ" = "upsilon",
                   "φ" = "phi", "χ" = "chi", "ψ" = "psi", "ω" = "omega")

# Function to replace Greek letters using the mapping
replace_greek = function(match) {
  greek_mapping[[match]]
}
```

Add front tags to match to output
```{r}
taxonomy_format = taxonomy %>%
  mutate(
    
    #Add taxon indicator
    superklass = sub("^", "p_", superklass),
    klass = sub("^", "c_", klass),
    species = sub("^", "s_", compound_name),
    
    #Remove alphanumeric characters
    superklass = str_replace_all(superklass, pattern = "[^_|[:alnum:]]", replacement = ""),
    klass = str_replace_all(klass, pattern = "[^_|[:alnum:]]", replacement = ""),
    species = str_replace_all(species, pattern = "[^_|[:alnum:]]", replacement = ""),
    
    #Replace Greek letters with their written out form
    superklass = str_replace_all(superklass, pattern = "[α-ωΑ-Ω]", replacement = replace_greek),
    klass = str_replace_all(klass, pattern = "[α-ωΑ-Ω]", replacement = replace_greek),
    species = str_replace_all(species, pattern = "[α-ωΑ-Ω]", replacement = replace_greek),
    
    #Lowercase
    superklass = str_to_lower(superklass),
    klass = str_to_lower(klass),
    species = str_to_lower(species),
    
    #replaces all "|" characters with underscores.
    superklass = gsub("\\|", "_", superklass),
    klass = gsub("\\|", "_", klass),
    species = gsub("\\|", "_", species)
  )
```

### Find the best match

Separate based on tags first.
```{r}
phylum = all_dataframes[grepl("^p_", all_dataframes$features), ]
class = all_dataframes[grepl("^c_", all_dataframes$features), ]
species = all_dataframes[grepl("^s_", all_dataframes$features), ]
```

Create taxonomy link
```{r}
taxonomy_links = left_join(taxonomy, taxonomy_format, by = 'compound_public_id')
```

Run the search, and link to it's formatted taxonomic name. 
Class
```{r}
best_matches_class =  find_best_matches(source_dataframe = class, target_dataframe = taxonomy_format, source_column = "features", target_column = "klass") 

class_format = taxonomy_links %>%
  select(c(klass.x, klass.y)) %>%
  rename('best_match' = 'klass.y',
         'formatted_name' = 'klass.x') %>%
  distinct(best_match, .keep_all = TRUE) %>%
  right_join(best_matches_class) %>%
  mutate(level = 'Class')
```

Phylum
```{r}
best_matches_phylum =  find_best_matches(source_dataframe = phylum, target_dataframe = taxonomy_format, source_column = "features", target_column = "superklass")

phylum_format = taxonomy_links %>%
  select(c(superklass.x, superklass.y)) %>%
  rename('best_match' = 'superklass.y',
         'formatted_name' = 'superklass.x') %>%
  distinct(best_match, .keep_all = TRUE) %>%
  right_join(best_matches_phylum) %>%
  mutate(level = 'Phylum')
```
Species or Compound Name
```{r}
best_matches_compound =  find_best_matches(species, taxonomy_format, "features", "species")

compound_format = taxonomy_links %>%
  select(c(compound_name.x, species)) %>%
  rename('best_match' = 'species',
         'formatted_name' = 'compound_name.x') %>%
  right_join(best_matches_compound) %>%
  mutate(level = 'Compound')
```

### Formatting

Combine all of the outputs
```{r}
# Add compound IDS
compound_ids = taxonomy %>% select(compound_public_id, compound_name)

# Combine all outputs, and merge with the Marker data
best_matches = full_join(phylum_format, class_format) %>%
  full_join(compound_format) %>%
  select(c(features, formatted_name, marker, level))%>%
  relocate(marker, 1) %>%
  left_join(compound_ids, by = c('formatted_name' = 'compound_name')) %>%
  
  #Assign taxonomic order
  mutate(level = factor(level, levels = c('Phylum', 'Class', 'Compound'))) %>%
  
  # Arrange by marker and taxonomi level
  arrange(marker, level)

write.csv(best_matches, 'output/HFE_pooled_selected_features_longform.csv', row.names = FALSE)
```


Condensed Form
```{r}
compiled_text = best_matches %>%
   mutate(label_with_ids = ifelse(level == 'Compound',
    paste(formatted_name, ' (', compound_public_id, ')', sep = ''), formatted_name)) %>%
  group_by(marker) %>%
  summarise(compiled_label = paste(label_with_ids, collapse = ', '))

write.csv(compiled_text, 'output/HFE_pooled_selected_features_compiled.csv', row.names = FALSE)
```

