---
title: "dietML metadata File preparation"
author: "Stephanie Wilson"
date: "October 2023"
output: html_document
---

#  dietML metadata file prep

__Required Input Files__.                          
  - *FL100_merged_variables.csv* - 
  - *metadata_CRP_subclinical.txt*
  - *metadata_fecal_subclinical.txt*
  
__Information__. 
This script organizes machine learning variable metadata files and organizes and copies them into the correct directories for dietML analyses.

__Output__. 
  - metadata files in the respective dietML folders

Load Packages and Data
```{r messages = FALSE, warnings = FALSE}
library(tidyverse)

merged = read.csv('data/FL100_merged_variables.csv') %>% 
  rename('subject_id' = 'UserName')
```

Function to write the same csv to multiple locations
```{r}
write_data_to_multiple_paths = function(target_locations, dataframe) {
  
  # Loop through each target location
  for (location in target_locations) {
    
    # Generate the full file path for the current location
    full_path = file.path(getwd(), location)
    
    # Write the dataframe to the current location
    write.csv(dataframe, file = full_path, row.names = FALSE)
  }
}
```


### 1) Select and format Machine Learning Variables

Select columns that are of interest to us for ML.
  - choosing to remove bristol_num (too many NAs)
  - choosing to remove hhincome (education can be just as an informative feature)
  - Oct 2023 - removing HEI Components due to overlap with Yasmine's paper
  - Oct 2023 - removing fecal pH and stool collection time
  
```{r}
ML_variables = merged %>%
  select(c(subject_id, Age:BMI, edu_level, avg_total_fiber, avg_total_PRO,
           #hei_asa24_totalveg:hei_asa24_totalscore, fecal_ph, diff_time_hrs
           hei_asa24_totalscore)) %>%
  filter(!is.na(avg_total_PRO)) %>%
  mutate(Sex = factor(Sex),
         edu_level = factor(edu_level))
```


### 2) SUBCLINICAL, Continuous Inflammation Outcomes

Subclinical Thresholds:
  - CRP < 10000 ng/mL
  - CAL < 100
  
```{r}
CRP_subclinical = read.delim('HFE/metadata_CRP_subclinical.txt', 
                             header = TRUE, sep = "\t", dec = ".") 
GI_subclinical = read.delim('HFE/metadata_fecal_subclinical.txt', 
                            header = TRUE, sep = "\t", dec = ".") 
```

```{r}
CRP_subclinical_meta = ML_variables %>%
  filter(subject_id %in% CRP_subclinical$subject_id)

GI_subclinical = ML_variables %>%
    filter(subject_id %in% GI_subclinical$subject_id) 
```

Create metadata paths
```{r}
CRP_sub_location = c("HFE/Subclinical/CRP_dietML/merged_blood_subclinical.csv")
CAL_sub_location = c("HFE/Subclinical/CAL_dietML/merged_GI_subclinical.csv")
MPO_sub_location = c("HFE/Subclinical/MPO_dietML/merged_GI_subclinical.csv")
NEO_sub_location = c("HFE/Subclinical/NEO_dietML/merged_GI_subclinical.csv")
```

```{r}
write_data_to_multiple_paths(dataframe = CRP_subclinical_meta, 
                             target_locations = CRP_sub_location)
write_data_to_multiple_paths(dataframe = GI_subclinical, target_locations = c(CAL_sub_location, MPO_sub_location, NEO_sub_location))
```

