---
title: "Understanding Food Sources - Compound Analysis"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Calculate Food Polyphenol Intakes

__Overview__
This code determines the average polyphenol intake for each unique food consumed by a participant.

__Inputs__
  - *PP_Raw.csv* - from A0_FilePrep_Exploration.Rmd
  - *FooDB_phenols_taxonomy.csv* - from A0A_Scrape_Taxonomy.Rmd
  - *Ingredient_Intake_Frequency.csv* - Output from A2_Polyphenol_Intake_Summary.Rmd
  
__Outputs__
  - *Avg_Polyphenol_Intake_ByFood.csv* - Avg Intake for each polyphenol from each unique food by each participants

```{r}
library(tidyverse)
```

```{r}
tax = read.csv('FooDB/FooDB_phenols_taxonomy.csv')

#Define compound outliers
removal_list = factor(c('Chinese tannin', 'alpha-Viniferin'))

# Load ASA - FooDB polyphenol data
PP_Raw = read.csv('data/FL100_PP_Raw.csv') %>%
  filter(!compound_name %in% removal_list) %>%
  left_join(tax) 

# Load food frequency data
frequency = read.csv('output/Ingredient_Intake_Frequency.csv') %>%
  select(-overall_avg_food_ppintake)
```



### Format Dietary Data

Averages across Foods by compound
```{r}
Intakes =  PP_Raw %>%
  
  #Group by Taxonomic Class
  group_by(UserName, RecallNo, Ingredient_code, compound_public_id) %>%
  
  #gets the sum of each compound for each participant's recall
  mutate(intake_mg = sum(pp_consumed, na.rm = TRUE)) %>% 
  select(c(UserName, food_id, food_name, RecallNo, Ingredient_code,
           Ingredient_description, Ingredient_consumed_g,
           compound_public_id, compound_name, klass, intake_mg)) %>%
  ungroup()%>%
  
  #Remove duplicates since we've summed each polyphenol per recall
  distinct(UserName, RecallNo, Ingredient_code, compound_public_id, .keep_all = TRUE)
```


```{r}
test = Intakes %>%
  filter(klass %in% c("Prenol lipds", "Phenylpropanoic acids")) %>%
  filter(intake_mg>0)
```


```{r}
Intakes_Ingredient = Intakes %>%
  
#Average polyphenol intake across recalls for each compound
  group_by(UserName, Ingredient_code, compound_public_id) %>%
  mutate(Avg_intake = mean(intake_mg, na.rm = TRUE)) %>%
  ungroup() %>%
  
  #Remove duplicates
  distinct(UserName, Ingredient_code, compound_public_id, .keep_all = TRUE) %>%
  select(c(UserName, RecallNo, food_id, food_name, Ingredient_code,
           Ingredient_description, compound_public_id,
           compound_name, klass, Avg_intake)) %>%
  filter(!is.na(Avg_intake))
```


```{r}
Intakes_Food = Intakes %>%
  
#Average polyphenol intake across recalls for each compound
  group_by(UserName, food_id, compound_public_id) %>%
  mutate(Avg_intake = mean(intake_mg, na.rm = TRUE)) %>%
  ungroup() %>%
  
  #Remove duplicates
  distinct(UserName, food_id, compound_public_id, .keep_all = TRUE) %>%
  select(c(UserName, RecallNo, food_id, food_name, compound_public_id,
           compound_name, klass, Avg_intake)) %>%
  filter(!is.na(Avg_intake))

write.csv(Intakes_Food, 'output/Avg_Polyphenol_Intake_ByFood.csv', row.names =FALSE)
```


### Test with one Compound
```{r}
# At Ingredient Level
Intakes_Ingredient %>%
  filter(compound_name == 'Epicatechin') %>%
  group_by(Ingredient_description, compound_public_id) %>%
  mutate(AVERAGE = mean(Avg_intake, na.rm = TRUE)) %>%
  distinct(Ingredient_description, .keep_all = TRUE) %>%
  select(c(food_name, Ingredient_description, klass, AVERAGE)) 

# At Food Level
Intakes_Food %>%
  filter(compound_name == 'Hyperin') %>%
  group_by(food_id, compound_public_id) %>%
  mutate(AVERAGE = mean(Avg_intake, na.rm = TRUE)) %>%
  distinct(food_id, .keep_all = TRUE) %>%
  select(c(food_id, food_name, AVERAGE)) %>%
  left_join(frequency)
```
