---
title: "diet ML - Feature Formatting"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Summarize and format feature names from dietML runs

This script pools output from multiple RDS files, which are a key output file from the dietML program built by Dr. Andrew Oliver. Each RDS file is a packaged R environment which contains numerous dataframes. This script:

1) Extract and formats feature names
2) Pull out SHAP importance data
3) Create Directionality Plot

__Inputs__. 
- RDS files from dietML output
- *rds_path_subclinical_names.csv*
- *HFE_pooled_selected_features_longform.csv*
- *metadata* - from ML2A_dietML_meta_Prep.Rmd
  
__Outputs__. 
- *dietML_top_features.csv*, top model features based on feature importance.
- Plots with model SHAP summaries.
  
```{r LOAD PACKAGES, message=FALSE}
library(tidyverse); library(rvest); library(wesanderson); library(xml2); library(stringr); library(Hmisc); library(ggpubr);library(grid)
```

Load Functions
```{r}
source('Functions/find_rds.R')
source('Functions/extract_rds_data.R')
```


Load Data
```{r}
#Load pooled taxaHFE output and metadata variable list
taxahfe_output = read.csv('output/HFE_pooled_selected_features_longform.csv')
metadata_full = str_to_lower(colnames(read.csv('HFE/Subclinical/NEO_dietML/merged_GI_subclinical.csv')))
rds_info = read.csv('output/rds_path_subclinical_names.csv')
```

### 1) Format Feature Names
```{r}
#Clean the compound data first
taxa = taxahfe_output %>%
  select(c(features, formatted_name, level)) %>%
  # Remove duplicate entries
  distinct(features, .keep_all = TRUE) %>%
  mutate(level2 = case_when(level == 'Phylum' ~ 'Superclass',
                            .default = level),
    feature_publish = ifelse(level2 %in% c('Class', 'Superclass'), 
                                  paste(formatted_name, ' (', level2, ')', sep = ''), 
                                  formatted_name)) %>%
  select(c(features, feature_publish)) %>%
  rename('feature' = 'features')
```

 
Format metadata names
```{r}
features = data.frame(feature = metadata_full, 
                      feature_publish = c('Subject', 'Age', 'Sex, Male', 'BMI',
                                          'Education',  'Avg Total Fiber', 
                                          'Avg Total Protein',  
                                          'ASA24 HEI Total Score')) %>%
  full_join(taxa) %>%
  mutate(feature = ifelse(feature == 'sex', "sex_male", feature))
```

### 2) Pull out SHAP importance data

Run Loop to pull out data from each of our RDS files
```{r warning = FALSE}
shap_data = list()

# Loop through each RDS file and extract the data
for (i in seq_along(rds_info$path)) {
  file_path = rds_info$path[i]
  extracted_data = extract_data_from_rds(file_path)
  shap_data[[i]] = extracted_data
}
```

Formatting
```{r}
# Use map2_dfr to iterate over each list and add the label to the MeanShaps dataframe
merged_MeanShaps = map2_dfr(shap_data, rds_info$label, add_label_to_MeanShaps) %>%
  #Remove the untransformed CRP Data and Keep the Transformed CRP Data
  filter(!label %in% c('Untransformed_Full_CRP', 'Untransformed_Mod_CRP')) %>%
  
  # Extract the text between the first and second underscore
  mutate(label = gsub("Untransformed|Transformed", "Continuous", label), 
         # label = gsub("LOHI", "Classifier", label), # no longer used
         label_model = str_extract(label, "_(.*?)_"),
         
         # Remove the underscores from the result
         label_model = gsub('_', '', label_model),
         
         #Extract the last three characters
         marker = substr(label, nchar(label) - 2, nchar(label)),
         
         #Simplify the phenolic data to not show full taxonomy
         feature_simple = ifelse(feature %in% metadata_full, feature, 
                                 gsub(".*_(.*_.*)", "\\1", feature)),
         #Pull out data on how the response was handled.
         response_type = sub("^(.*?)_.*", "\\1", label),
         
         #Indicate directionality
         Direction = ifelse(Avg_SHAP > 0, 'Positive', 'Negative'),
         axis_label = sub("_[^_]*$", "", label),
         
         #Relevel markers
         marker = factor(marker, levels = c("CRP", "CAL", "MPO", "NEO"))) %>%
           
  left_join(features, by = 'feature') %>%
  relocate(feature_publish, .after= 'feature') %>%
  select(-c(marker, label_model)) %>%
  rename('marker' = 'response_type')
```

```{r}
top_features = merged_MeanShaps %>%
  #For each model outcome combination
  group_by(marker) %>%
  #Extract top ten features based off Importance
  top_n(n = 5, wt = Importance) %>%
  arrange(desc(Importance))

write.csv(top_features, 'output/dietML_top_features.csv', row.names = FALSE)
```


### 3) Directionality Plot

Model plots in publication are represented by beeswarm plots. 


```{r}
PLOT = ggplot(top_features, aes(x = marker, y = reorder(feature_publish, Importance), fill = Direction)) + 
  labs(x = "", y = '', 
       subtitle = '')+
  geom_point(aes(colour = factor(Direction)), size = 2) +
  theme_bw() + 
  coord_fixed(ratio = 0.3) + 
  guides(color = guide_legend(title = "Direction")) + 
  scale_color_manual(values = c("#CB7A5C", "#5785C1")) + 
  theme(legend.position = 'bottom',
        legend.margin = margin(0, 0, 0, 0),
      legend.spacing.x = unit(0, "mm"),
      legend.spacing.y = unit(0, "mm"),
        plot.margin = margin(c(0, 0, 0, 0)))

PLOT
```


