---
title: "dietML - RDS Paths and Performance"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Create RDS Paths and Extract Performance Metrics

This script pools output from multiple RDS files, which are a key output file from the dietML program built by Dr. Andrew Oliver. Each RDS file is a packaged R environment which contains numerous dataframes. This script:

1) Isolates the pathways for each of the RDS files created by one run of dietML and 
2) Extracts performance metrics from random forest models

__Inputs__. 
- *dietML RDS files* - from dietML output
- *metadata* - from ML2A_dietML_meta_Prep.Rmd
- *HFE_pooled_selected_features_longform.csv* - Output from ML1B_Pooled_taxaHFE_features.Rmd
  
__Outputs__. 
- *rds_path_subclinical_names.csv*
- *dietML_subclinical_performance_metrics.csv*
  
```{r LOAD PACKAGES, message=FALSE}
library(tidyverse); library(rvest); library(wesanderson); library(xml2); library(stringr); library(Hmisc); library(ggpubr);library(grid)
```

Load Functions
```{r}
source('Functions/find_rds.R')
source('Functions/extract_rds_data.R')
source('Functions/extract_performance_metrics.R')
```


Load Data
```{r}
#Load pooled taxaHFE output and metadata variable list
taxahfe_output = read.csv('output/HFE_pooled_selected_features_longform.csv')
metadata_full = str_to_lower(colnames(read.csv('HFE/Subclinical/NEO_dietML/merged_GI_subclinical.csv')))
```

### 1) Format Feature Names
```{r}
#Clean the compound data first
taxa = taxahfe_output %>%
  select(c(features, formatted_name, level)) %>%
  # Remove duplicate entries
  distinct(features, .keep_all = TRUE) %>%
  mutate(level2 = case_when(level == 'Phylum' ~ 'Superclass',
                            .default = level),
    feature_publish = ifelse(level2 %in% c('Class', 'Superclass'), 
                                  paste(formatted_name, ' (', level2, ')', sep = ''), 
                                  formatted_name)) %>%
  select(c(features, feature_publish)) %>%
  rename('feature' = 'features')
```

 
Format metadata names
```{r}
features = data.frame(feature = metadata_full, 
                      feature_publish = c('Subject', 'Age', 'Sex, Male', 'BMI',
                                          'Education',  'Avg Total Fiber', 
                                          'Avg Total Protein',  
                                          'ASA24 HEI Total Score')) %>%
  full_join(taxa) %>%
  mutate(feature = ifelse(feature == 'sex', "sex_male", feature))
```

### 2) Load Path Information

Get the paths of all rds files
```{r}
rds_files = find_rds_files(getwd())
rds_files_df = as.data.frame(str_remove(rds_files, '/Users/stephanie.wilson/Desktop/SYNC/Scripts/FL100_Polyphenol_Analysis/')) %>%
  dplyr::rename('path' = 1) %>%
  filter(str_detect(path, "Subclinical")) %>%
  slice_head(n=4)
```

Extract Informative Key Labels from the RDS Paths 
```{r}
tags = as.data.frame(str_match(rds_files_df$path, 
                               #Extract text between 1 and 2nd, 2nd and 3rd /
                               # extract the text before the underscore
                               "/([^/]+)/([^/]+)/([^/]+)_")[, 2:4]) %>%
  # Rename the variables to be more descriptive
  dplyr::rename('response_type' = 1,
                'model' = 2,
                'marker' = 3) %>%
  mutate(model = ifelse(model == 'Modifiable', 'Mod', model))

#combine the dataframes
rds_info = cbind(rds_files_df, tags) %>%
  mutate(label = paste(response_type, model, marker, sep = '_'))

write.csv(rds_info, 'output/rds_path_subclinical_names.csv', row.names = FALSE)
```



### 3) Performance metrics
```{r warning = FALSE}
metrics_data = list()

# Loop through each RDS file and extract the data
for (i in seq_along(rds_info$path)) {
  file_path = rds_info$path[i]
  extracted_data = extract_performance(file_path)
  metrics_data[[i]] = extracted_data
}

# Format
all_metrics = bind_rows(metrics_data) %>%
  rename('metric' =1 ,
         'estimator' = 2, 
         'estimate' = 3, 
         'config' = 4, 
         'label' = 5) %>%
  left_join(rds_info) %>%
  select(-path) 

#Export
write.csv(all_metrics, 'output/dietML_subclinical_performance_metrics.csv', row.names = FALSE)
```
