---
title: "Random Forest Beeswarm Plots"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Create dietML Beeswarms of top five features

This script pools SHAP data from multiple RDS files and creates beeswarm plots. Plots show top five features in the prediction of subclinical inflammation outcomes.

__Inputs__. 
- *RDS files* from dietML output
- *rds_path_subclinical_names.csv*, from ML3B_dietML_RDSpaths_Performance.Rmd
- *dietML_top_features.csv*, from ML3A_dietML_Feature_Formatting.Rmd 
  
__Outputs__. 
- Plots with SHAP summary plots.
  

```{r message = FALSE, warning = FALSE}
library(tidyverse); library(shapviz); library(ggpubr); library(cowplot)
source('Functions/create_beeswarm_RDS.R')
# image folder
imagepath = '/Users/stephanie.wilson/Desktop/SYNC/Scripts/FL100_Polyphenol_Analysis/images/'
```


```{r}
rds = read.csv('output/rds_path_subclinical_names.csv') %>%
  mutate(marker = substr(response_type, 1, 3),
         label = paste0('Subclinical ', marker))

# Split the list into model types
# classification = rds %>% filter(response_type == 'LOHI')
regression = rds 
```


```{r}
# Load model features and their publishing-friendly names
top_features = read.csv('output/dietML_top_features.csv') %>%
  select(c(feature, feature_publish)) %>%
  distinct(feature, .keep_all = TRUE)
```

### 1) Regression  Beeswarm Plots
```{r warning = FALSE}
regression_list = list()

# Loop through paths in your dataframe
for (i in seq_along(regression$path)) {
  path = regression$path[i]
  feature_list = top_features
  opt = regression$marker[i]
  
  plot_result = create_beeswarm(path, feature_list)
  
  # Generate a unique name for each plot based on the opt values
  unique_name = paste0("plot_", opt, sep = "")
  
  # Store the plot_result in the list with the unique name
  regression_list[[unique_name]] = plot_result
}
```

### 2) Arrange Plots
```{r}
# Arrange plots
regression_plots = ggarrange(regression_list[[2]], regression_list[[1]],
          regression_list[[3]], regression_list[[4]] + labs(x = 'SHAP value'),
          ncol = 1, nrow = 4, 
          common.legend = TRUE, legend = 'right', align = "v",
          labels = c('A) CRP', 'B) CAL', 'C) MPO', 'D) NEO'), 
          hjust = 0, vjust = c(1.5, 0, 0, 0))
regression_plots
```

```{r}
#Save
ggsave('Regression_Beeswarms.jpg', plot = regression_plots, path = imagepath, height = 7, width =6, units = 'in', dpi = 300)
ggsave('Regression_Beeswarms.pdf', plot = regression_plots, path = imagepath, height = 7, width =6, units = 'in')
```

