---
title: "Polyphenol Feature Food Sources"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Identify Top 3 Food Sources for Top Random Forest Polyphenol Features

__Summary__. 
This script helps determine the top food contributors to top polyphenol features. Food contributors and related model information is provided in a supplementary manuscript table. 

__Inputs__. 
  - *FL100_PP_DailyAveraged.csv*, Output from A0B_Exploration.Rmd
  - *FooDB_phenols_taxonomy.csv*, Output from A0A_Scrape_Taxonomy.Rmd
  - *Ingredient_Intake_Frequency.csv*, Output from A2_Polyphenol_Intake_Summary.Rmd
  - *Avg_Polyphenol_Intake_ByFood.csv*, Output from ML3_dietML_Polyphenol_Food_Intakes.Rmd
  - *dietML_top_features.csv*, Output from ML3A_dietML_Feature_Formatting.Rmd
  
__Outputs__. 
  - *dietML_topPP_features_foodsources.csv*, Has top polyphenol features and the top three food sources for each. Features most relevant to our inflammation models additionally have directionality with outcomes noted.

```{r}
library(tidyverse); library(gt); library(ggrepel)
```


Load Data
```{r}
# What are our top 10 features
features = read.csv('output/dietML_top_features.csv') 
tax = read.csv('FooDB/FooDB_phenols_taxonomy.csv')

# Avg Polyphenol Intake Across Participants By food_id
Intakes_Food = read.csv('output/Avg_Polyphenol_Intake_ByFood.csv')

# Load food frequency data
frequency = read.csv('output/Ingredient_Intake_Frequency.csv') %>%
  select(-overall_avg_food_ppintake)

# Polyphenol content for each compound in foods, averaged across each participant's recalls.
content = read.csv('data/FL100_PP_DailyAveraged.csv') %>%
  select(UserName, food_id:compound_name, mean_daily_intake_mg) %>%
  left_join(tax) 
```

### 1) Generate Food table Data

#### 1a) ALL taxonomic Higher Groupings
```{r}
class = features %>%
  filter(grepl("^c_", feature)) %>%
  mutate(taxa = sub("\\s*\\(Class\\)$", "", feature_publish))

class_foods = content %>%
  filter(klass %in% class$taxa) %>%
  group_by(food_id, klass) %>%
  mutate(Class_avg = mean(mean_daily_intake_mg, na.rm = TRUE)) %>%
  select(food_id, food_name, klass, Class_avg)%>%
  distinct(food_id, .keep_all = TRUE) %>%
  left_join(frequency) %>%
  ungroup() %>%
  
  # Remove NaNs and zeros
  filter(complete.cases(Class_avg)) %>%
  filter(Class_avg>0) %>%
  # Keep more frequent foods
  group_by(klass) %>%
  filter(food_freq >= median(food_freq)) %>%
  arrange(desc(Class_avg)) %>%
  slice_head(n = 3) %>%
  
  # Format for table
  summarise(compiled_foods = paste(food_name, collapse = ', ')) %>%
  mutate(taxon = paste(klass, '(Class)', sep =' ')) %>%
  select(-klass) %>%
  relocate(taxon, 1)
```

```{r}
superclass =features %>%
  filter(grepl("^p_", feature)) %>%
  mutate(taxa = sub("\\s*\\(Superclass\\)$", "", feature_publish))

sp_foods = content %>%
  filter(superklass %in% superclass$taxa) %>%
  group_by(food_id, superklass) %>%
  mutate(superklass_avg = mean(mean_daily_intake_mg, na.rm = TRUE)) %>%
  select(food_id, food_name, superklass, superklass_avg)%>%
  distinct(food_id, .keep_all = TRUE) %>%
  left_join(frequency)  %>%
  ungroup() %>%
   # Remove NaNs and zeros
  filter(complete.cases(superklass_avg)) %>%
  filter(superklass_avg>0) %>%
  # Keep more frequent foods
  filter(food_freq >= median(food_freq)) %>%
  arrange(desc(superklass_avg)) %>%
  group_by(superklass) %>%
  slice_head(n = 3) %>%

  # Format for table
  summarise(compiled_foods = paste(food_name, collapse = ', ')) %>%
  mutate(taxon = paste(superklass, '(Superclass)', sep =' ')) %>%
  select(-superklass) %>%
  relocate(taxon, 1)
```

#### 1b) ALL COMPOUNDS in our ML model
```{r}
ALL = Intakes_Food %>%

  left_join(frequency) %>%
  
  # Get the cohort average for each food
  group_by(food_id, compound_public_id) %>%
  mutate(AVERAGE = mean(Avg_intake, na.rm = TRUE)) %>%
  distinct(food_id, .keep_all = TRUE) %>%
  select(compound_name, food_name, food_freq, klass, AVERAGE) %>%
  arrange(desc(AVERAGE)) %>%
  ungroup()
```

```{r}
compound_foods = ALL %>%
  filter(!AVERAGE == 0) %>%
  group_by(compound_name) %>%
  filter(food_freq >= median(food_freq)) %>%
  slice_head(n = 3) %>%
  summarise(compiled_foods = paste(food_name, collapse = ', ')) %>%
  rename(taxon = 1)
```

#### 1c) Merge higher order taxa with compounds
```{r}
merged = full_join(sp_foods, class_foods) %>%
  full_join(compound_foods)
```


### 2) Compiled Markers

Create Dataframe of polyphenol features and their associated directionality with the inflammation outcome.
```{r}
feature_minimal = features %>%
  
  select(feature_publish, marker, Direction) %>%
  filter(feature_publish %in% merged$taxon) %>%
  mutate(Direction = ifelse(Direction=="Positive", '(+)', '(-)'),
         marker_label = paste(marker, Direction, sep = ' ')) %>%
  select(-c(marker, Direction)) %>%
  group_by(feature_publish) %>%
  summarise(compiled_marker = paste(marker_label, collapse = ', ')) %>%
  rename('taxon' = 1)
```

Create Table
```{r}
#Add food intake information with feature information
table_data = merged %>%
  
  left_join(feature_minimal) %>%
  rename('Taxon or Compound' = 1,
         'Associated Markers (Direction)' = 3,
         'Common Food Sources' = 2)

write.csv(table_data, 'output/dietML_topPP_features_foodsources.csv', row.names =FALSE)

gt_tbl = gt(table_data)
gt_tbl

gtsave(filename = 'images/ML_polyphenol_food_sources.png', 
       data = gt_tbl)
```
