---
title: "Random Forest Regression Plots"
author: "Stephanie Wilson"
date: "November 2023"
output: html_document
---

# Create regression plots for the top 5 random forest model features. 
Resulting plots are included as supplementary manuscript material. These are in addition to model beeswarm plots.

__Inputs__. 
  - *RDS files* from dietML runs
  - *rds_path_subclinical_names.csv*, from ML3B_dietML_RDSpaths_Performance.Rmd
  - *dietML_top_features.csv*, from ML3A_dietML_Feature_Formatting.Rmd
  
__Outputs__. 
  - Regression Plots images

```{r warning = FALSE, message = FALSE}
library(tidyverse); library(ggpubr)
```

Load Top 5 dietML features and their original rds files
```{r}
features = read.csv('output/dietML_top_features.csv') %>%
  select(c(feature, feature_publish, marker))

rows_to_rename = c(3, 11, 14, 15, 18)
# New names for the specified rows
new_names = c("Avg Total Protein (g)", "ASA24 HEI Total Score", 
              "Avg Total Fiber (g)", "BMI (kg/m^2)", "Age (years)")

features_unique = features %>%
  select(-marker) %>%
  distinct(feature, .keep_all = TRUE) %>%
  mutate(feature_publish = case_when(
    row_number() %in% rows_to_rename ~ new_names[match(row_number(), rows_to_rename)],
    TRUE ~ feature_publish),
    feature_publish = ifelse(feature_publish %in% new_names, feature_publish, paste(feature_publish, "(mg)")))

rds = read.csv('output/rds_path_subclinical_names.csv') %>%
  mutate(marker = substr(response_type, 1, 3),
         label = paste0('Subclinical ', marker))
```
`


### 1) Extract RDS Input data

Create a function to extract only the input data
```{r warning = FALSE}
extract_input_from_rds = function(path) {

  load(path) # Load the RDS file
  input = input # Extract the required dataframes
  
  return(data.frame(input))
}
```

### 2) Pull and format the data for each marker
```{r warning = FALSE}
CRP = extract_input_from_rds(rds$path[2]) %>% 
  select(c(subject_id, label, 
           one_of(features %>% filter(marker=='CRP') %>% select(feature) %>% pull()))) %>% 
  rename('CRP_boxcox' = 'label') %>%
  pivot_longer(cols =3:7, names_to = 'feature', values_to = 'value') %>%
  left_join(features_unique, by = 'feature')

CAL = extract_input_from_rds(rds$path[1]) %>% 
  select(c(subject_id, label, 
           one_of(features %>% filter(marker=='CAL') %>% select(feature) %>% pull()))) %>% 
  rename('Calprotectin' = 'label') %>%
  pivot_longer(cols =3:7, names_to = 'feature', values_to = 'value') %>%
  left_join(features_unique, by = 'feature')

MPO = extract_input_from_rds(rds$path[3]) %>% 
  select(c(subject_id, label, 
           one_of(features %>% filter(marker=='MPO') %>% select(feature) %>% pull()))) %>% 
  rename('Myeloperoxidase' = 'label') %>%
  pivot_longer(cols =3:7, names_to = 'feature', values_to = 'value') %>%
  left_join(features_unique, by = 'feature')

NEO = extract_input_from_rds(rds$path[4]) %>% 
  select(c(subject_id, label, 
           one_of(features %>% filter(marker=='NEO') %>% select(feature) %>% pull()))) %>% 
  rename('Neopterin' = 'label') %>%
  pivot_longer(cols =3:7, names_to = 'feature', values_to = 'value') %>%
  left_join(features_unique, by = 'feature')
```

### 3) Create regression plots

```{r}
CRP_plot = ggscatter(CRP, x = "value", y = "CRP_boxcox", add = "reg.line") +
  facet_wrap(~str_wrap(feature_publish, width =25), scales = 'free_x', ncol=5) +
  coord_cartesian(ylim = c(min(CRP$CRP_boxcox), max(CRP$CRP_boxcox)*1.6))+
  stat_cor(method = 'spearman', p.accuracy = 0.001,
           label.x.npc = 'left', label.y = max(CRP$CRP_boxcox)*1.4, size =3) + 
  labs(y = 'CRP (Box-Cox)', x = '')+ 
    theme(plot.margin = unit(c(0,0,0,0.5), "cm")) 

CAL_plot = ggscatter(CAL, x = "value", y = "Calprotectin", add = "reg.line") +
  facet_wrap(~str_wrap(feature_publish, width =23), scales = 'free_x', ncol=5) +
  coord_cartesian(ylim = c(min(CAL$Calprotectin), max(CAL$Calprotectin)*1.3))+
  stat_cor(method = 'spearman', p.accuracy = 0.001,
           label.x.npc = 'left', label.y = max(CAL$Calprotectin)*1.2, size =3) + 
  labs(y = 'CAL (ng/mL)', x = '')+ 
    theme(plot.margin = unit(c(0,0,0,0.5), "cm"))

MPO_plot = ggscatter(MPO, x = "value", y = "Myeloperoxidase", add = "reg.line") +
  facet_wrap(~feature_publish, scales = 'free_x', ncol=5) +
   coord_cartesian(ylim = c(min(MPO$Myeloperoxidase), max(MPO$Myeloperoxidase)*1.3))+
  stat_cor(method = 'spearman', p.accuracy = 0.001,
           label.x.npc = 'left', label.y = max(MPO$Myeloperoxidase)*1.2, size =3) +
  labs(y = 'MPO (ng/mL)', x = '')+ 
    theme(plot.margin = unit(c(0,0,0,0.5), "cm"))

NEO_plot = ggscatter(NEO, x = "value", y = "Neopterin", add = "reg.line") +
  facet_wrap(~str_wrap(feature_publish, width = 25), scales = 'free_x', ncol=5) +
 coord_cartesian(ylim = c(min(NEO$Neopterin), max(NEO$Neopterin)*1.3))+
  stat_cor(method = 'spearman', p.accuracy = 0.001,
           label.x.npc = 'left', label.y = max(NEO$Neopterin)*1.2, size =3) +
  labs(y = 'NEO (ng/g)', x = '') + 
    theme(plot.margin = unit(c(0,0,0,0.5), "cm"))

combined_CRP_CAL = ggarrange(CRP_plot, CAL_plot, nrow = 2, labels = c("A)", "B)"))
combined_MPO_NEO = ggarrange(MPO_plot, NEO_plot, nrow = 2, labels = c("A)", "B)"))
```


```{r}
#Main Text
ggsave(plot = combined_CRP_CAL, filename = 'images/dietML_CRP_CAL_correlations.png', width = 9, height = 4, dpi = 300)
ggsave(plot = combined_CRP_CAL, filename = 'images/dietML_CRP_CAL_correlations.pdf', width = 9, height = 4)

# Supplemental
ggsave(plot = combined_MPO_NEO, filename = 'images/dietML_MPO_NEO_correlations.png', width = 9, height = 4, dpi = 300)
ggsave(plot = combined_MPO_NEO, filename = 'images/dietML_MPO_NEO_correlations.pdf', width = 9, height = 4)
```


